---
title: "I do it for fun"
author: "Charles Kwame Appiah"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    code_download: true
    theme: sandstone
    toc: true
  pdf_document:
    toc: true
---
# Theme
"bootstrap", "cerulean", "cosmo", "darkly", "flatly", "journal", "lumen", "paper", "readable", "sandstone", "simplex", "spacelab", "united", "yeti"

# Data Types
Data types are used to represent different types of data in R. R has several built-in data types, such as `numeric`, `character`, `integer`, `complex`, and `logical`. Data types are used to store different types of data, such as numbers, text, and logical values.


```{r}
x <- "Hello Charles, I hope you're working hard to become who you are destined to be by God's purpose? "
x
class(x) # checking the class of variable x
```

```{r}
y <- pi^2 # calculating the square of pi
y
class(y) # checking the class of variable y
```

```{r}
z <- 15L # assigning 15 to variable
z
class(z) # checking the class of variable z
```

```{r}
a <- (5 + 2i)^2 # calculating the square of (5 + 2i)
a
class(a) # checking the class of variable a
```

```{r}
l <- TRUE # assigning TRUE to variable l
class(l) # checking the class of variable l
```

```{r}
x <- list(age=c(10,21,22), weight=c(30,33,32)) # creating a list x
x
names(x) # Calling the names of the list x

length(x) # checking the length of the list x
```

```{r}
kx <- data.frame(age=c(11,14,22), weight=c(30,33,32)) # creating a dataframe xk 
kx
```

# Vector and List

```{r}
d <- c("Charles Kwame Appiah is my name") # creating a vector
d
class(d) # checking the class of variable d
length(d) # checking the length of variable d
nchar(d) # checking the number of characters in variable d
```

```{r}
f<- c(1,3,4,6,7) # creating a vector
f
class(f) # checking the class of variable f
```
Integers in R are stored as numeric data type. To create an integer in R, you can use the `L` suffix or the `as.integer()` function.
```{r}
fo<- c(1L,3L,4L,6L,7L) # creating a vector
fo
class(fo) # checking the class of variable fo
```

# Vectors
Vectors can be created with the `c()` function. The `c()` function can take multiple arguments and combine them into a single vector. The `c()` function can also be used to combine multiple vectors into a single vector.
```{r}
# Initialization
x <- vector(mode = "logical", length = 5) # creating a vector with 5 logical elements
x
class(x)

x[1:3] <- TRUE # indexing the first to third element with TRUE
x
```
# Logical
Logical vectors are created with the `c()` function. Logical vectors can be used to store TRUE or FALSE values. Logical vectors are used in R to represent binary data.Binary data is data that can take on one of two values, such as TRUE or FALSE.

```{r}
s <- c(TRUE, FALSE,TRUE, 1) # creating a vector
s
as.logical(s) # converting the vector to logical
```
# List
List is a collection of multiple objects. The objects can be of different classes. Lists are created with the `list()` function

```{r}
q <- list("Hello World",2015L, TRUE, 32.1) # creating a list
q
class(q[[2]]) # Checking the class of list 2
class(q[[4]]) # Checking the class of list 4
class(q[[1]]) # Checking the class of list 1
```

# Matrix
Matrix is a two-dimensional array. It is created with the `matrix()` function. The `matrix()` function takes a vector as input and reshapes it into a matrix. The `nrow` argument specifies the number of rows in the matrix, and the `ncol` argument specifies the number of columns in the matrix.

```{r}
mat <- c(2,4,5,7) # creating a vector
dim(mat) <- c(2,2) # creating a matrix with 2 rows and 2 columns
mat

mar = c(2,4,5,7,8,9) # creating a vector
dim(mar) <- c(3,2) # creating a matrix with 3 rows and 2 columns
mar
```

```{r}
temp <- c(3,4,5,5.6,6,7) # creating a vector
mati <- matrix(temp, nrow = 2, ncol = 3,byrow = TRUE) # creating a matrix with 2 rows and 3 columns
mati

temp <- c(3,4,5,5.6,6,7, 8, 9, 10) # creating a vector
mato <- matrix(temp, nrow=3, ncol=3,byrow=TRUE) # creating a matrix with 3 rows and 3 columns
mato
```

```{r}
# Default byrow = FALSE
temp <- c(3,4,5,5.6,6,7) # creating a vector
mati <- matrix(temp, nrow=2, ncol=3,byrow=FALSE) # creating a matrix with 2 rows and 3 columns
mati
```

```{r}
t1 <- c(23, 55) # creating a vector
t2 <- c(34, 45) # creating a vector

# By rows
rbind(t1,t2) # binding them by their rows
```

```{r}
t3 <- c(32, 50)
t4 <- c(43, 54)

# By columns
cbind(t3,t4) # binding them by their columns
```

# Factors and Nominal Data
Factors are used to represent categorical data in R. Factors are created with the `factor()` function. The `factor()` function takes a vector as input and converts it into a factor. Factors are used to represent nominal data, which is data that has no inherent order.

```{r}
factor <- c("Yes","No","No","Yes") # creating a vector
factor # use to encode vectors as factors

f <- factor(c("Yes","No","No","Yes"), levels= c("Yes","No")) # creating a factor
f
```

# Missing Values
Missing values are represented by the `NA` value in R. The `is.na()` function can be used to check if a value is missing. The `is.na()` function returns `TRUE` if the value is missing and `FALSE` otherwise.

```{r}
x <- NA # Missing number
x

is.na(x) # checking if it is a missing number 
```

```{r}
u <- 0/0 # Missing number
u
class(u) # checking for the class of variable u
```

# Dataframe
Data frames are used to store tabular data in R. Data frames are created with the `data.frame()` function. The `data.frame()` function takes vectors as input and combines them into a data frame. Data frames are used to store data in a structured format.

```{r}
c <- c("Charles","Richmond","Nicholas") # creating a vector
d <- c(12, 23, 45) # creating a vector
s <- c(FALSE,TRUE,TRUE) # creating a vector

dfr <- data.frame(Username = c,Age = d, Adult = s) # creating a dataframe
dfr 
```

```{r}
# First Row
dfr[1,] # accessing row 1 of the dataframe

# First Column
dfr[,1] # accessing column 1 [Username] of the dataframe

# Age Column
dfr$Age

# Username Column
dfr$Username

# Adult Column
dfr$Adult
```

# Reading Data From File
Reading data from a file is a common task in data analysis. R provides functions to read data from various file formats, such as CSV, Excel, and text files. The `read.csv()` function is used to read data from a CSV file. The `read.csv()` function takes the path to the CSV file as input and returns a data frame. Sample data is provided in the `Training r.csv` file.

```{r }
### Importing Datasets
library(readxl) # Importing the readxl package

dat <- read.csv("C:/Users/HP/Desktop/Data Science/Machine learning/Training r.csv") # reading the dataset
#print(dat, na.rm = TRUE)
head(dat) # displaying the first few rows of the dataset


data <- read_xlsx("C:/Users/HP/Desktop/Data Science/Pandas/Copy of V1- UN Data on Refugees (AiCE __ Dataset).xlsx") # reading the dataset
 head(data) # displaying the first few rows of the dataset
```

# Sequence Creation
Sequence is a collection of numbers in a specific order. Sequences can be created in R using the `:` operator, the `seq()` function, and the `rep()` function. The `:` operator is used to create a sequence of numbers from a starting value to an ending value. The `seq()` function is used to create a sequence of numbers with a specified start, end, and step size. The `rep()` function is used to repeat a sequence of numbers a specified number of times.

```{r}
# Sequence
v <- (10:20) # creating a sequence
v # start from 10 and end at 20

w <- (-5:9)
w # start from -5 and end at 9
```

```{r}
qw <- seq(2,34,2) # creating a sequence
qw # start from 2 and end at 34 with a moving step of 2

iqw <- seq(2,34,length=6) # creating a sequence 
iqw # start from  2 and end at 34 with a length of 6
```

```{r}
repe <- rep(1:4,4) # creating a sequence 
repe # repeat from 1 to 4, 4 times

eq <- rep("Hello Ann", 5) # creating a sequence
eq # repeat from Hello Ann, 5 times

we <- seq(1,15, 2) # creating a sequence 
we  # start from 1 and end at 15 with a step of 2

we[1:5] # slicing from index 1 to 5

class(we) # checking the class of we
```

```{r}
fo <- list("Hello","Hi","Hey") # creating a list
fo
fo[c(1,2)] # for several elements
fo[c(1,2,3)] # for several elements
fo[[2]] # for only one element

class(fo[[3]]) # checking the class of the third element
```

```{r}
wi <- list(age=c(12,23,45), height=c(12.3,45.4, 34.5)) # creating a list
wi
class(wi) # checking the class of wi
```

```{r}
woo <- data.frame(age = c(12,23,45), height = c(12.3,45.4, 34.5)) # creating a dataframe
woo
class(woo) # checking the class of woo

wi$age # accessing only the age list
wi$height # accessing only the height list

wi[["age"]] # accessing only the age list
wi[['h',exact = FALSE]] # partial matching

class(wi$age) # checking the class of age list
class(wi$height) # checking the class of height list
```

```{r}
wr <- matrix(1:9, nrow = 3, ncol = 3, by = TRUE) # creating a matrix with 3 rows and 3 columns
wr # creating a matrix with 3 rows and 3 columns
class(wr) # checking the class of wr

class(wr[1,1]) # checking the class of row 1 column 1

class(wr[1,1, drop=FALSE]) # checking the class of row 1 column 1
```

```{r}
ch <- c(1:9,NA,NA,NA) # creating a vector
ch
i<- is.na(ch) # checking for missing numbers
i
ch[!i] # removing missing numbers
```

# Vectorization
Vectorization is a technique used in `R` to perform operations on vectors without using loops. Vectorization allows you to perform operations on vectors more efficiently than using loops. Vectorization is a key concept in `R` programming and is used to write efficient and concise code.

```{r}
### Vectorization
ew <- rnorm(1000) # creating a vector

er <- rnorm(1000) # creating a vector

cv <- vector(mode="numeric", length=1000) # creating a vector

### Iteration
start <- proc.time() # starting the timer
for (i in 1:1000){ 
cv[i] <- ew[i] + er[i] # adding the two vectors
}
proc.time()-start # initiating the timer


start <- proc.time() # starting the timer
cv <- ew + er # adding the two vectors
proc.time()-start # # initiating the timer
```

# Control Structures
Control structures are used to control the flow of a program. Control structures allow you to execute different code blocks based on conditions. Control structures include if-else statements, for loops, while loops, and repeat loops.

```{r}
x <- 20 # assigning 20 to x

if (x < 0) {
print("Negative!") # checking if x is negative
}else if (x < 10) {
print("Positive, less than 10!") # checking if x is less than 10
}else {
print("Number greater than 10!") # checking if x is greater than 10
}

x <- -20
if (x < 0) {
print("Negative!") # checking if x is negative
}else if (x < 10) {
print("Positive, less than 10!") # checking if x is less than 10
}else {
print("Number greater than 10!") # checking if x is greater than 10
}

x <- 6 # assigning 6 to x
if (x < 0) {
print("Negative!") # checking if x is negative
}else if (x < 10) {
print("Positive, less than 10!") # checking if x is less than 10
}else {
print("Number greater than 10!") # checking if x is greater than 10
}
```
# For Loop
`For loop` is used to execute a block of code for  a specified number of times. `For loops` are used when you know the number of iterations in advance. `For loops` are used to iterate over a sequence of values.

```{r}
for (i in 1:100){
cat(i)
cat(" ") # inserting spaces between the numbers
}

letters # lower case 

LETTERS # upper case

class(letters) # 

for (x in letters){
cat(x)
cat(" ") # inserting spaces between the letters
}
```
# While loop
`While loops` are used to execute a block of code as long as a condition is true. `While loops` are used when you do not know the number of iterations in advance. `While loops` are used to execute a block of code as long as a condition is true. `While loops` are used when you do not know the number of iterations in advance.

```{r}
x <- -1 # assigning -1 to x
while (x < 5){
print(x) # 
x <- x+1
}

x<- 1
repeat{
print(x)
if (x > 7){
break
}
x <- x+1
}
```

```{r}
for (i in 1:100){
# Over ride the first 20 iterations
if (i <= 20){
next 
}
}

# Functions
myPrinter <- function(x){
  for (i in seq(x)){
       print("Hello, Charles")
  }
}
myPrinter(3) 

volume <- function(x=3, y=3, z=3){
  print(x*y*z)
}
volume(y=3,z=5,x=11)

volume()

myPrinter <- function(..., mes){
print(sum(...))
print(mes)
}
myPrinter (3, 5, 11, mes= "Hi! Richmond")

#ls() ### displaying objects stored in R currently
```

# Iterated Functions 
Iterated function systems (IFS) are a method of constructing fractals; the resulting fractals are often self-similar. IFS fractals are more related to set theory than fractal geometry. They were introduced in 1988 by Michael Barnsley using the concept of affine transformations.

## lapply
lapply is used to apply a function to each element of a list. lapply returns a list of the same length as the input list, where each element is the result of applying the function to the corresponding element of the input list.

```{r}
# lapply
str(lapply)

x <- list(a=rnorm(10), b=rnorm(20), c=rnorm(30))
lapply(x, mean) # checking the mean of x
lapply(x, var)# checking the variance of x
lapply(x, sd) # checking the standard deviation of x
```
## sapply
sapply is used to apply a function to each element of a list and simplify the result. sapply returns a vector or matrix instead of a list.

```{r}
# sapply
str(sapply) # checking the structure of sapply

xi <- list(a=rnorm(10), b=rnorm(20), c=rnorm(30))
sapply(xi, mean) # checking the mean of xi
sapply(xi, var) # checking the variance of xi
sapply(xi, sd) # checking the standard deviation of xi
```
## Split
split is used to split a data frame or vector into groups. split returns a list of vectors, where each vector contains the elements of the input data frame or vector that belong to a particular group.

```{r}
# Split
dat <- data.frame(subject = 1:6,age = c(15, 17, 16,20,21,23),
adult = c(FALSE,FALSE,FALSE,TRUE,TRUE,TRUE))
s <- split(dat, dat$adult)
s # split them according to True and False


sapply(s, function(x){
mean(x[["age"]])
})

sapply(s, function(x){
var(x[["age"]])
})

sapply(s, function(x){
sd(x[["age"]])
})
```

## tapply 
tapply is used to apply a function to each group of a vector. tapply returns a vector or matrix, where each element is the result of applying the function to the corresponding group of the input vector.

```{r}
# tapply
str(tapply)

x <- c(rnorm(10),rnorm(10),rnorm(10),rnorm(10))
f <- gl(4, 10)
f
tapply(x, f, mean)
tapply(x, f, var)
tapply(x, f, sd)
```

```{r}
### Help 
#?c
#?vector
#?sapply
#?lapply
#?tapply
```

# Types, Quality and Data preprocessing
Types, quality and data preprocessing are important steps in data analysis. Types refer to the data types of the variables in the dataset. Quality refers to the quality of the data, such as missing values and outliers. Data preprocessing refers to the steps taken to clean and prepare the data for analysis. 

```{r}
wi

# finding each column maximum
m <- sapply(wi,max)
m

# finding each column minimum
n <- sapply(wi,min)
n
```
# Regularization
Regularization is a technique used to prevent overfitting in machine learning models. Regularization adds a penalty term to the loss function to prevent the model from fitting the training data too closely. Regularization is used to improve the generalization of the model.r
```{r}
# Regularization ith range [0,1]
wi$age <- ( (wi$age - n[1])/(m[1] - n[1]))*(1 - 0) + 0
wi$height <- ( (wi$height - n[2])/(m[2] - n[2]))*(1 - 0)  

wi
```

# DPLYR AND TIDYR PACKAGES
`Dplyr` is a package for data manipulation in R. `Dplyr` provides a set of functions for manipulating data frames, such as selecting columns, filtering rows, and summarizing data. `Tidyr` is a package for data tidying in R. `Tidyr` provides functions for reshaping data frames, such as gathering columns into rows and spreading rows into columns.

```{r}
library(dplyr) # Importing the dplyr package
data(airquality) # loading the airquality dataset
class(airquality) # checking the class of airquality
airquality <- tibble(airquality) # converting airquality to a tibble
class(airquality) # checking the class of airquality
airquality # displaying the airquality dataset
select(airquality, Ozone, Solar.R, Day) # selecting the Ozone, Solar.R and Day columns

select(airquality, -(Wind:Month)) # offsetting Wind and Month from the airquality dataset
filter(airquality, Month > 5, Month < 9, Day < 3) # values in Month greater than 5 and less than 9, Day values less than 3

filter(airquality, Day==1 | Day == 2) # Values of Day = 1 or 2
arrange(airquality,Ozone, desc(Solar.R))
mutate(airquality, Temp.C = round((Temp - 32) * 5/9)) # creating a new column for Temp.C

# Removing rows with missing values on the Ozone and Solar.R features
airquality <- filter(airquality, !is.na(Ozone),!is.na(Solar.R))
airquality 
### print(airquality, n=143)
```

```{r}
# grouping by month
by_month <- group_by (airquality, Month) 
by_month

# Finding the minimum, average and maximum value per Month
summarize (by_month, min(Ozone), mean(Ozone), max(Ozone))
```

```{r}
#install.packages ("tidyr") #library (tidyr)

#dat #gather(dat, sex, count, -subject)

#dat \<- gather(dat, sex, class, count, -subject) #dat

#dat #separate(dat, sex, class, c("sex", "class"))

#dat #dat \<- gather(dat, lesson, grade, lesson1:lesson4, na.rm = TRUE)
#dat

#dat \<- spread(dat, quarter, grade) #dat

#mutate(dat, lesson = extract_numeric(lesson))
```

# Statistical Summary and Visualization
`Statistical summary and visualization` are important steps in data analysis. Statistical summary provides a summary of the data, such as the mean, median, and standard deviation. Visualization provides a visual representation of the data, such as bar charts and histograms.

```{r}
# Mean
internet_usage = c(22,0, 7,12,5, 33, 14, 8, 0, 9)

#internet_usage
mean(internet_usage) # finding the mean of internet_usage

net_usage = c(22,0,7,12,5,NA,33,14,8,NA,0,9)

# net_usage
mean(na.omit(net_usage)) # finding the mean of net_usage with missing numbers 
```

```{r}
# Median
median(net_usage, na.rm = TRUE)

# Minimum, Maximum and Range
A = c(49,33,63,48,54,62,52,64,71,68)
min(A)
max(A)
which.min(A)
which.max(A)
```

```{r}
print(max(A) - min(A))
range(A)
print(range(A)[2] - range(A)[1]) 
summary(A)
```

# Percentile Values
Percentile values are used to divide a dataset into equal parts. The `quantile()` function is used to calculate percentile values in R. The `quantile()` function takes a vector as input and returns the percentile value for the specified quantile.

```{r}
X = c(3,4,5,6,7,8,10,10,11,12,14,14,14,15,16,17,21,25,27,32)
quantile(X,0.80)
quantile(X,0.50,type = 7)
quantile(X,0.25,type = 7)
quantile(X,0.75,type = 7)
median(X)
summary(X)
sd(X)
# cv(X)
```
## Interquartile Range
The interquartile range (IQR) is a measure of statistical dispersion, or how spread out the values in a dataset are. The IQR is calculated as the difference between the 75th percentile (Q3) and the 25th percentile (Q1) of the dataset. The IQR is used to identify outliers in a dataset.

```{r}
# Interquartile Range
irq = function(X) (quantile(X,0.75) - quantile(X,0.25))
irq(X)
```
## Variance and Standard Deviation
Variance and standard deviation are measures of statistical dispersion. Variance is the average of the squared differences between each value in the dataset and the mean of the dataset. Standard deviation is the square root of the variance. Variance and standard deviation are used to measure the spread of the values in a dataset.

```{r warning=FALSE}
# Variance and Standard Deviation
course = c(6, 2, 1, 9, 17, 4, 3, 2, 1, 5, 11 ,4, 3, 1, 2, 2, 5, 4, 3, 6)
1 / course 
cf = c(course, 0, course)
cf

vw <- 2 * course + cf + 1
vw
```

```{r}
var(course)
sum((course - mean(course)) ^ 2 / (length(course) - 1))

sd(course)
sqrt(var(course))
std = function(x) sqrt(var(x))
std(course)
sqrt(course)
sum(course)
prod(course)
sort(course)
order(course)
sqrt(-14 + 9i)
```

## Coefficient of Variation
The coefficient of variation (CV) is a measure of relative variability. The CV is calculated as the standard deviation divided by the mean of the dataset. The CV is used to compare the variability of datasets with different units of measurement.

```{r}
cv = function(x) ( sd(x) / mean(x) )
cv(course)
```
# Visualization of Qualitative Data
`Visualization of qualitative data` is an important step in data analysis. Qualitative data is data that is categorical or non-numeric. Visualization of qualitative data can help identify patterns and trends in the data. Bar charts, pie charts, and contingency matrices are commonly used to visualize qualitative data.

```{r}
mo = c("car","car","bus", "metro","metro","car","metro","metro","foot","car","foot","bus","bus","metro","metro","car","car","car","metro","car") # dataset of employee's 
mo

table(mo) # creating a table of mo
prop.table(table(mo)) # creating a table of mo
data.frame(mo)# creating a dataframe of mo
```

## Bar Charts
`Bar charts` are used to visualize the frequency of categorical data. Bar charts are used to compare the frequency of different categories in a dataset. Bar charts are used to visualize the distribution of categorical data.

```{r}
barplot(table(mo)) # creating a bar chart of mo

barplot(prop.table(table(mo))) # creating a bar chart of mo
```

## Pie Chart
`Pie charts` are used to visualize the proportion of different categories in a dataset. Pie charts are used to compare the proportion of different categories in a dataset. Pie charts are used to visualize the distribution of categorical data.

```{r}
pie(table(mo), col = c ("red", "green", "blue", "black")) # creating a pie chart of mo

pie(prop.table(table(mo)), col = c("purple", "green", "red", "blue")) # creating a pie chart of mo
```

# Contingency Matrix
`A contingency matrix` is used to visualize the relationship between two categorical variables. A contingency matrix is a two-dimensional table that shows the frequency of each combination of categories in the two variables. A contingency matrix is used to visualize the relationship between two categorical variables.

```{r}
g = c(rep("Male",8), rep("Female",12)) # creating a vector
g

mg = table(mo,g) # creating a contingency matrix of mo and g
mg

gm = data.frame(mo,g) # creating a dataframe of mo and g
gm


margin.table(mg, 1) # creating a margin table of mg
margin.table(mg,2) # creating a margin table of mg
prop.table(mg)     # creating a table of mg
prop.table(mg,1)   # creating a table of mg
prop.table(mg,2)   # creating a table of mg
```


## Stacked Bar Charts and Grouped Bar Charts
`Stacked bar charts` are used to visualize the relationship between two categorical variables. Stacked bar charts are used to compare the frequency of different categories in one variable for each category in another variable. Stacked bar charts are used to visualize the relationship between two categorical variables.

```{r}
barplot(mg, col = c("red","blue", "pink","green")) # creating a bar chart of mg


barplot(mg,xlim=c(0,2), xlab="Gender", length=levels(mo), col = c("red","blue", "pink","green")) # creating a bar chart of mg

barplot(mg,xlim=c(0,2), xlab="Transportation", length=levels(g), col = c("red","blue", "pink","green")) # creating a bar chart of mg

barplot(prop.table(mg,1), width=0.25, xlim=c(0,3), ylim=c(0,1), xlab="Gender", legend=levels(mo), beside=T, col = c("red","blue", "pink","green")) # creating a bar chart of mg

mg = table(g,mo) # creating a contingency matrix of g and mo
barplot(prop.table(mg,2), width=0.25, xlim=c(0,3), ylim=c(0,1), xlab="Transportation", legend=levels(g), beside=T, col= c("red", "green")) # creating a bar chart of mg
```


## Histograms
`Histograms` are used to visualize the distribution of numerical data. Histograms are used to show the frequency of different values in a dataset. Histograms are used to visualize the distribution of numerical data.

```{r}
xo = c(10, 10, 5, 9, 7, 6,8,6,5,8, 10, 7, 7,8, 5, 6,4,7,9,7, 4,8, 10,10, 7,4,9,5,8,9) # creating a vector
table(xo) # creating a table of xo
data.frame(xo) # creating a dataframe of xo

prop.table(table(xo)) # creating a table of xo
#?prop.table


hist(xo, col = c("red","blue", "pink","green")) # creating a histogram of xo
hist(xo, nclass = 3, col = c("red","blue", "pink","green")) # creating a histogram of xo
```

```{r}
wo = c (1950, 2090, 2700, 3350, 4200, 3720, 4400, 2980, 3850, 4550, 3050, 2350, 1850, 2820, 3670, 2950, 3750, 1850, 2420, 3150, 3000, 3470, 3920, 3100, 2400, 2900, 2650, 3450, 3650, 4020, 4450, 3120, 3660, 3070, 3550, 2020, 3500, 2500, 3780, 3940, 3540, 2800, 4450, 1950, 3020, 2800, 3500, 1480, 4495,2850, 3100, 2250,3300, 4100, 3220, 3600,2130, 4020, 4075) # creating a vector

hist(wo, col = c("red","blue", "pink","green")) # creating a histogram of wo
hist(wo, nclass=4, col = c("red","blue", "pink","green")) # creating a histogram of wo
hist(wo, breaks= seq(from = 1000, to=5000, by=300), col = c("red","blue", "pink","green")) # creating a histogram of wo
hist(wo, probability=T, col = c("red","blue", "pink","green")) # creating a histogram of wo
rug(jitter(wo)) # creating a histogram of wo
```

## Frequency Polygon
`Frequency polygons` are used to visualize the distribution of numerical data. Frequency polygons are used to show the frequency of different values in a dataset. Frequency polygons are used to visualize the distribution of numerical data.

```{r}
temp = hist(wo, col = c("red","blue", "purple","green")) # creating a histogram of wo
temp

lines(c(min(temp$breaks), (temp$mids),max(temp$breaks)), c(0,temp$counts,0),type="l") # creating a frequency polygon of wo
```
## Boxplot
`Boxplots` are used to visualize the distribution of numerical data. `Boxplots` are used to show the median, quartiles, and outliers in a dataset. `Boxplots` are used to visualize the distribution of numerical data.

```{r}
boxplot(wo, horizontal = T, col = c("red")) # creating a boxplot of wo
boxplot(wo, vertical = T, col = c("red")) # creating a boxplot of wo
fivenum(wo) # summary  of wo
summary(wo) # summary of wo

#?fivenum
```

```{r}
w1 = c(1950, 2090, 2700, 3350, 4200, 3720, 4400, 2980, 3850, 4550, 3050, 2350,1850, 2820, 3670, 2950, 3750, 1850, 2420,3150, 3000, 3470, 3920, 3100, 2400, 2900, 2650, 3450, 3650, 4020) # creating a vector of w1

fivenum(w1) # summary of w1

w2 = c(4450, 3120, 3660, 3070, 3550, 2020, 3500, 2500, 3780, 3940, 3340, 2800, 2850, 4450, 1950,3020,2800,3500,1480,4495,2850,3100,2250,3300,4100,3220,3600,2130,4020,4075) # creating a vector of w2

fivenum(w2) # summary of w2

boxplot(w1, w2, names = c("Sample 1", "Sample 2"), col = c("purple", "yellow" )) # creating a boxplot of w1 and w2
```

# Classification and Prediction
`Classification and prediction` are important tasks in data analysis. Classification is the process of categorizing data into different classes. Prediction is the process of predicting the value of a target variable based on the values of other variables. `Classification` and `prediction` are used to make decisions based on data.

```{r}
computeCost <- function(X, y, th){
m <- length(y)
return(1/(2*m) * sum((X%*%th - y)^2))
}

computeCost(5,6,7) # computing the cost of 5,6,7

grad_desc <- function(X, y, theta,alpha, lambda, num_iters){
m <- length(y)
F_history <- c(rep(0, num_iters)) # creating a vector of zeros

for (iter in c(1:num_iters)){
  temp <- vector()
  temp <- theta * (1 - ((alpha*lambda)/m)) - alpha*(1/m) *   (t(X) %*% (X %*% theta - y))
  theta <- temp
  F_history[iter] <- computeCost(X, y, theta) 
}
print(F_history[num_iters])
return(list("theta" = theta, "F_history" = F_history))
}

grad_desc(2,3,5,0.1,7.5,2) # computing the output
```
# Clustering
Clustering is a technique used to group similar data points together. Clustering is used to identify patterns and relationships in data. Clustering is used to group data points into clusters based on their similarity. Clustering is used to analyze and explore data.

`kmeans` is a popular clustering algorithm that is used to group data points into k clusters. kmeans is an unsupervised learning algorithm that is used to identify patterns and relationships in data. kmeans is used to group data points into clusters based on their similarity.

```{r}
iris_new <- iris # assigning iris to iris_new
### iris_new
iris_new$Species <- NULL # assigning the species class to null
kc <- kmeans(iris_new, 3) # creating a kmeans of iris_new
table(iris$Species, kc$cluster) # creating a table of iris and kc

plot(iris_new[c("Sepal.Length", "Sepal.Width")], col = kc$cluster) # visualizing the  iris datasets 
points(kc$centers[,c("Sepal.Length","Sepal.Width")], col = 1:3, pch = 8, cex = 2)


plot(iris_new[c("Petal.Length", "Petal.Width")], col = kc$cluster) # visualizing the iris dataset 
points(kc$centers[,c("Petal.Length","Petal.Width")], col = 1:3, pch = 8, cex = 2)
```

```{r}
data(iris) # importing the iris dataset
set.seed(500) # setting the seed to 500 to avoid randomness
idx <- sample(1:dim(iris)[1], 40) # sampling the iris dataset
iris_Sample <- iris[idx,] # assigning the sampled iris dataset to iris_Sample
iris_Sample$Species <- NULL # assigning the species class to null
hc <- hclust(dist(iris_Sample), method = "single") # creating a hclust of iris_Sample

plot(hc, hang = -1, labels = iris$Species[idx], xlab = "Clusters") # visualizing the iris dataset
rect.hclust(hc, 3) # creating a rect.hclust of 3
hc <- hclust(dist(iris_Sample),method = "complete") # creating a hclust of iris_Sample
plot(hc, hang = -1, labels = iris$Species[idx], xlab = "Clusters") # visualing the iris dataset
rect.hclust(hc, 3) # creating a rect.hclust of 3
```

```{r}
data(iris)
set.seed(500)
idx <- sample(1:dim(iris)[1], 50)
iris_Sample <- iris[idx,]
iris_Sample$Species <- NULL
hc <- hclust(dist(iris_Sample), method = "single")
plot(hc, hang = -1, labels = iris$Species[idx], xlab = "Clusters")
rect.hclust(hc, 3)
hc <- hclust(dist(iris_Sample),method = "complete")
plot(hc, hang = -1, labels = iris$Species[idx], xlab = "Clusters")
rect.hclust(hc, 3)
```

```{r}
tr = trees # assigning trees to tr

as.data.frame(tr) # converting tr to a dataframe

plot(tr[c("Volume","Height")], col = "red") # visualizing the trees dataset

plot(tr[c("Girth","Height")], col = "blue" ) # visualizing the trees dataset

plot(tr[c("Volume","Girth")], col = "black" ) # visualizing the trees dataset
```

```{r}
ca = cars # assigning cars to ca 

as.data.frame(ca) # converting ca to a dataframe

plot(ca[c("speed","dist")], type = "l") # visualizing the cars dataset
```

```{r}
pre = pressure # assigning pressure to pre

as.data.frame(pre) # converting pre to a dataframe

plot(pre[c("temperature","pressure")], type="l") # visualizing the pressure dataset
```

```{r}
cab = CO2 # assigning CO2 to cab

as.data.frame(cab) # converting cab to a dataframe

plot(cab[c("conc","uptake")]) # visualizing the CO2 dataset
```

```{r}
oran = Orange # assigning Orange to oran

as.data.frame(oran) # converting oran to a dataframe

plot(oran[c("Tree","circumference")]) # visualizing the Orange dataset

plot(oran[c("Tree","age")]) # visualizing the Orange dataset 

plot(oran[c("age","circumference")]) # visualizing the Orange dataset
```

# Mining of frequent itemsets and Association Rules
Mining of frequent `itemsets` and association rules is an important task in data analysis. Frequent `itemsets` are sets of items that occur together frequently in a dataset. Association rules are rules that describe the relationships between items in a dataset. Mining of frequent `itemsets` and association rules is used to identify patterns and relationships in data.

```{r warning=FALSE}
library(arules) # importing the arules package

db <- list(c("A", "B", "D", "E"), c("B", "C", "E"), c("A", "B", "D", "E"), c("A", "B", "C", "E"), c("A", "B", "C", "D", "E"), c("B", "C", "D")) # creating a list of items

frequent <- apriori(db, parameter = list(supp = 0.5, conf = 1, target="frequent itemsets")) # creating a frequent itemset of db

inspect(frequent) # inspecting the frequent itemset

cl <- apriori(db, parameter = list(supp = 0.5, conf = 1, target = "closed")) # creating a closed itemset of db

inspect(cl) # inspecting the closed itemset

mx <- apriori(db, parameter=list(supp=0.5, conf=1, target="maximal"))
# creating a maximal itemset of db

inspect(mx) # inspecting the maximal itemset

rules <- apriori(db, parameter=list(supp=0.5, conf=1, target="rules")) # creating a rule of db

inspect(rules) # inspecting the rule
```

```{r}
data(Adult) # importing the Adult dataset

dim(Adult) # checking the dimensions of Adult

inspect(apriori(Adult, parameter=list(supp = 0.75))) # inspecting the Adult dataset

inspect(apriori(Adult, parameter = list(supp = 0.75), appearance = list(rhs = "capital-gain=None", default = "lhs"))) # inspecting the Adult dataset

```
