---
title: "Fun stuffs"
author: "Charles"
date: "2025-03-24"
output:
  pdf_document: default
  html_document: default
  always_allow_html: true
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## **Chapter One**

### Simple Expressions

```{r}
2 + 4 # Addition

4 / 2 # Division

3 * 4 # Multiplication

45 - 13 # Subtraction

4L / 5L # Integer Division
```

```{r}
4 %% 3 # for integer division with remainder

2 ^ 5 # Exponential

3 ** 4 # Indices

9 ** 3 # Indices
```

```{r}
(s = "Hello Charles, I hope you are working smart to achieve your dream?")


e <- 34 # Assigning variable (e) to  value (34)
e
nchar(e) # number of characters

3 -> r # Assigning variable (r) to  value (3)
r
```

```{r}
(u <- "Nana Kwame") # parentheses form will print an output
nchar(u) # number of characters

q = seq( 1,50,2) # Start from 1 and end at 50 with a step of 2
q

q[23] # checking for index of the sequence

q[20] # checking for index of the sequence

q[2:9] # checking form index 2 to index 9

q ** 2 - r
```

```{r}
v <- c("A" = 1, "B" = 2, "C" = 3) # Creating keys and values (dictionaries)
v

names(v) <- c("x", "y", "z") # Assigning new keys to the values
v

v["x"] # checking the key x

v["y"] # checking the key y
```

```{r}
(x <- 1:4) # parentheses form will print an output

(y <- 4:7) # parentheses form will print an output

x - y 
```

```{r}
square <- function(x) x**2
square(1:4)

cube = function(x) x**3
cube(1:9)

square_and_subtract <- function(x, y) return(x ** 2 - y)
square_and_subtract(30,30)

sum(1:4) # sum 1 to 4

1 : 4 # output 1 to 4

sqrt(1:4) # squaer root of 1 to 4
```

```{r}
average <- function(x) {
 n <- length(x)
 sum(x) / n
}
average(1:9)

#mean(1:9)

mean <- function(x) sum(x) / length(x)

mean(2:24)
```

### **Control loops** {#sec-controls-loops}

```{r}
# if statement

if (3 > 2) "true"

if (2 < 3) "false" else "true"

if (2 > 3) "bar" else "baz"

if (5 > 4) "Charles" else "Nana"

if (2 > 3) {
 x <- "bar"
}else "baz"
```

```{r}
# ifesle statement

x <- 1:5
ifelse(x > 3, "bar", "baz")

c <- 2 : 7
ifelse(c <- 6, "Nana", "Kwame")

even_square <- function(x) {
 ifelse (x %% 2 == 0, x ** 2, x)
}
even_square(1:9)
```

```{r}
# for loop

xo <- 1:5 # Assigning values (1 to 5) to variable xo
total <- 0
for (element in xo) total <- total + element
total
xo
```

```{r}
x <- 1:5
total <- 0
for (index in seq_along(x)) {
 element <- x[index]
 total <- total + element
}
total
```

```{r}
# While loop

x = -1:-5 # Assigning values (-1 to - 5) to variable x
total <- 9
index <- 1
while (index <= length(x)) {
 element <- x[index]
 index <- index + 1
 total <- total + element
}
total

x

summary(x)
```

### Factors

```{r}
(f <- factor(c("small", "small", "medium",
 "large", "small", "large")))

ordered(f)

summary(f)
```

```{r}
ff <- factor(c("small", "small", "medium",
 "large", "small", "large"),
 levels = c("small", "medium", "large")) # changing the levels of ff

ff

ordered(ff)

summary(ff)
```

```{r}
f <- factor(c("small", "small", "medium",
 "large", "small", "large"),
 levels = c("medium", "small", "large"),
 ordered = TRUE)
f
summary(f)
```

```{r}
v <- 1:26
names(v) <- LETTERS[1:26] # Extracting the name of the letters with numbers
v

(ffo <- factor(LETTERS[1:4]))
v[ffo] # Assigning values to the letters

(ff <- factor(LETTERS[1:4], levels = rev(LETTERS[1:4]))) 
v[ff]

as.numeric(ff) # Numeric form
as.vector(ff) # String form
v[as.vector(ff)]
```

### Data Frames

```{r}
df <- data.frame(a = 1:4, b = letters[1:4]) # Creating a data frame
df

df[1,1] # row one , column one

df[1,2] # row one , column two

df[2,1] # row two , column one

df[2,2] # row two , column two

df[1,] # All row one values

df[,1] # All column one values

df[,"a"] # using the column name

df$b # selecting only column name
```

```{r}
df2 <- data.frame(a = 5:7, b = letters[5:7])
rbind(df, df2) # Combining them using their rows

df3 <- data.frame(c = 5:8, d = letters[5:8])
cbind(df, df3) # Combining them using their columns

df4 = data.frame(Random = rnorm(26), Alphabets =  letters)
df4
```

### Missing Values

```{r}
(NA + 5)

is.na(NA) # checking for a missing number

is.na(4) # checking if 4 is a missing number

v <- c(1:100,NA)
sum(v)
sum(v, na.rm = TRUE)
```

### **Using R Packages**

#### Data Pipeline

```{r data}
library(magrittr)

subsample_rows <- function(d, n) {
 rows <- sample(nrow(d), n)
 d[rows,]
}

d <- data.frame(x = rnorm(100), y = rnorm(100))

d %>% subsample_rows(n = 100)
```

```{r}
do <- data.frame(x = rnorm(100), y = rnorm(100))

do %>% lm(y ~ x, data = .)
```

```{r}
so <- data.frame(x = 1:10, y = 11:20)

so %>% lm(y ~ x, data = .)
```

```{r}
#?lm()

rnorm(4) %>% data.frame(x = ., is_negative = . < 0)

rnorm(4) %>% data.frame(x = ., y = abs(.))

rnorm(4) %>% data.frame(x = sin(.), y = cos(.))

#sin(1.03130453)
#cos(1.03130453)
```

```{r}
f <- function(x, y) {
 y <- sum(cos(x),sin(y)) 
 z <- sum(sin(x),cos(y))
 z
}

f(2,4)

f(8,9)

f(4, 7)
```

```{r}
plot_and_fit <- function(d) {
 plot(y ~ x, data = d)
 abline(lm(y ~ x, data = d))
}

x <- rnorm(20)
y <- x + rnorm(20)
data.frame(x, y) %>% plot_and_fit

```

```{r}
data.frame(x, y) %>% (function(d) {
 plot(y ~ x, data = d)
 abline(lm(y ~ x, data = d))
})
```

```{r}
data.frame(x, y) %>% {
 plot(y ~ x, data = .)
 abline(lm(y ~ x, data = .))
}
```

```{r}
d <- data.frame(x = rnorm(100), y = 4 + rnorm(100))

#d %>% {data.frame(mean_x = mean(.$x), mean_y = mean(.$y))}

d %$% data.frame(mean_x = mean(x), mean_y = mean(y))

di <- data.frame(x = rnorm(10), y = rnorm(10))

di %T>% plot(y ~ x, data = .) %>% lm(y ~ x, data = .)
```

## **Chapter Three**

### Data Manipulation

```{r warning=FALSE}
# importing packages
library(datasets)
library(ggplot2)
library(magrittr)
#library(help = "datasets")

data(cars) # inintiating cars dataset
#head(cars)
cars %>% head(6) ### First 6 rows of the cars datasets
cars %>% tail(6) ### Last 6 rows of the cars datasets
cars %>% summary ### Summary of the cars datasets
cars %>% qplot(speed, dist, data = ., col = "brown")
```

```{r}
#install.packages("mlbench")
library(mlbench)
#library(help = "mlbench")

data(iris) # initiating the iris dataset
#head(iris)
iris %>% head(6) # First 6 rows of the iris dataset
iris %>% summary # Summary of the iris datasets
iris %>% qplot(Sepal.Length, Sepal.Width, data = ., color = Species)
iris %>% qplot(Petal.Length, Petal.Width, data = ., color = Species)
```

**Breast Cancer Dataset**

```{r}
#?read.table
data(BreastCancer)
BreastCancer %>% head()
data_url = "http://tinyurl.com/kw4xtts"
lines <- readLines(data_url)
lines[1:5]
```

```{r}
raw_breast_cancer = read.csv("C:/Users/HP/Desktop/Data Science/Machine learning/breast+cancer+wisconsin+original/breast-cancer-wisconsin.data")
raw_breast_cancer  |> head(6)
```

```{r}
raw_breast_cancer <- read.csv(data_url)
raw_breast_cancer %>% head()
```

```{r}
raw_breast_cancer <- read.csv(data_url, header = FALSE)
raw_breast_cancer %>% head()
```

```{r}
names(raw_breast_cancer) <- names(BreastCancer)
raw_breast_cancer %>% head()
```

```{r}
raw_breast_cancer <- read.csv(data_url, header = FALSE,
 col.names = names(BreastCancer))
raw_breast_cancer %>% head()
```

```{r}
formatted_breast_cancer <- raw_breast_cancer

#mapped <- formatted_breast_cancer$Class %>% map_class
#mapped %>% table

map_class <- function(x) {
 ifelse(x == 2, "bening", "malignant")
}
mapped <- formatted_breast_cancer$Class %>% map_class
mapped %>% table

formatted_breast_cancer$Class %>% unique

#dict <- c("2" = "benign", "4" = "malignant")
#map_class <- function(x) dict[as.character(x)]

#mapped <- formatted_breast_cancer$Class %>% map_class
#mapped %>% table
```

```{r}
mapped[1:20]

#mapped %<>% unname
#mapped[1:20]

read = read.csv(data_url, header = FALSE,
 col.names = names(BreastCancer)) ->
 raw_breast_cancer ->
 formatted_breast_cancer
dict <- c("2" = "benign", "4" = "malignant")
map_class <- function(x) dict[as.character(x)]
formatted_breast_cancer$Class <-
 formatted_breast_cancer$Class %>%
 map_class %>%
 unname %>%
 factor(levels = c("benign", "malignant"))

read

raw_breast_cancer$Class %>%
 { dict <- c("2" = "benign", "4" = "malignant")
 dict[as.character(.)]
 } %>%
 unname %>%
 factor(levels = c("benign", "malignant")) %>%
 table
```

### Saving a file and loading a saved file

```{r}
formatted_breast_cancer %>%
save(file = "C:/Users/HP/Desktop/Data Science/Learning R Language/formatted-breast-cancer.rda")

  
load("C:/Users/HP/Desktop/Data Science/Learning R Language/formatted-breast-cancer.rda") 
```

### Boston Housing Datasets

```{r}
#data(DNA)
#str(DNA)


data(BostonHousing)
str(BostonHousing)

boston <- "http://tinyurl.com/zq2u8vx"
```

```{r}
#data(BostonHousing2)
#str(BostonHousing2)
```

```{r}
#boston_housing <- read.table(boston)
#str(boston_housing)

col_classes <- rep("numeric", length(BostonHousing))
col_classes[which("chas" == names(BostonHousing))] <- "factor"

#boston_housing <- read.table(boston,col.names = names(BostonHousing),colClasses = col_classes)
#str(boston_housing)
```

### Readr Package

```{r BreastCancer}

library(readr)

raw_breast_cancer <- read_csv("C:/Users/HP/Desktop/Data Science/Machine learning/breast+cancer+wisconsin+original/breast-cancer-wisconsin.data",show_col_types = "TRUE", col_names = names(BreastCancer))
raw_breast_cancer %>% as.data.frame(head())
```

```{r}
raw_breast_cancer <- read_csv("C:/Users/HP/Desktop/Data Science/Machine learning/breast+cancer+wisconsin+original/breast-cancer-wisconsin.data",
 col_names = names(BreastCancer), show_col_types = "TRUE")
raw_breast_cancer %>% as.data.frame(head())
```

### Manipulating Data with dplyr

### Selection

```{r warning=FALSE}
library(tidyverse)

iris %>% tibble::as_tibble()
iris %>% tbl_df %>% select(Petal.Width,Species) %>% head()
iris %>% tbl_df %>% select(Sepal.Width,Species) %>% head()
iris %>% tbl_df %>% select(Species,Petal.Length) %>% head()
iris %>% tbl_df %>% select(Species,Sepal.Length) %>% head()
```

```{r}
iris %>% tbl_df %>%
select(Sepal.Width, Petal.Width, Species) %>% head(10) ## Select Sepal.Width, Petal.Width and Species from iris
```

```{r Excluding Petal.Width }
# Select from Sepal.Width to Petal.length
iris %>% tbl_df %>% select(Sepal.Length:Petal.Length) %>% head(10) 
```

```{r Petal only}
# Select Petal.Length and Width only
iris %>% tbl_df %>% select(starts_with("Petal")) %>% head(10) 
```

```{r Length Only, warning=FALSE}
# Select Petal.Length and Sepal.Length only
iris %>% tbl_df %>% select(ends_with("Length")) %>% head(10)
```

```{r etal Only}
# Select columns containing etal only
iris %>% tbl_df %>% select(contains("etal")) %>% head(10)
```

```{r}
iris |> select(Sepal.Length : Petal.Width)
```

```{r al Only}
# Select columns containing al only
iris %>% tbl_df %>% select(contains("al")) %>% head(10)
```

```{r Without Petal}
# Select columns that matches with t
iris %>% tbl_df %>% select(matches(".t.")) %>% head(10)

#  selection without Petal
iris %>% tbl_df %>% select(-starts_with("Petal")) %>% head(10) 
```

## Mutate

```{r Petal. (Width + Length)}
# Creating a new column with Petal width and length
iris %>% tbl_df %>% mutate(Petal.Width.plus.Length = Petal.Width + Petal.Length) %>% select(Species, Petal.Width.plus.Length) %>% head(100)
```

```{r Sepal. (Width + Length)}
# Creating a new column with Sepal width and length
iris %>% tbl_df %>% mutate(Sepal.Width.plus.Length = Sepal.Width + Sepal.Length) %>% select(Species, Sepal.Width.plus.Length) %>% head(30)
```

```{r Petal,Sepal.(Width + Length)}
# Creating a new column with Petal.Sepal width and length
iris %>% tbl_df %>% mutate(Petal.Width.plus.Length = Petal.Width + Petal.Length,Sepal.Width.plus.Length = Sepal.Width + Sepal.Length) %>% select(Species,Petal.Width.plus.Length, Sepal.Width.plus.Length) %>% head(30)
```

## Transmute

```{r Mutate.Select}
# Creating new column with transmute 
iris %>% tbl_df %>% transmute(Species, Petal.Width.plus.Length = Petal.Width + Petal.Length) %>% head()
```

```{r}
# Creating new column with transmute
iris %>% tbl_df %>% transmute (Species, Sepal.Length.plus.Sepal.Width = Sepal.Width + Sepal.Length) %>% head()

iris |> transmute (Species, Sepal.Length.plus.Sepal.Width = Sepal.Width + Sepal.Length) %>% head(20)
```

## Arrange

```{r Small to large}
# arrange from smallest to largest in Sepal.Length column
iris %>% tbl_df %>% arrange(Sepal.Length) %>% head() 

```

## 

```{r Small.Large}
# arrange from smallest to largest in Petal.Width column
iris %>% tbl_df %>% arrange(Petal.Width) %>% head()

```

```{r Large.Small}
# arrange from largest to smallest in Sepal.Length column
iris %>% tbl_df %>% arrange(desc(Sepal.Length)) %>% head() 
```

## Filter

```{r Values >= 5}
# Filter values >= 5 in the Sepal.Length
iris %>% tbl_df %>% filter(Sepal.Length >= 5) %>% head()
```

```{r Values <=1.4}
# Filter values <= 1.4 in the Peta.Length
iris %>% tbl_df %>% filter(Petal.Length <= 1.4) %>% head()
```

```{r Values > 5 and virginica}
# Filter values > 5 in the Sepal.Length and is Virginca species
iris %>% tbl_df %>% filter(Sepal.Length > 5 & Species == "virginica")  %>% select(Species, Sepal.Length) %>% head() 
```

## Groupby

```{r Species grouping}
# Grouping by the Species column
iris %>% tbl_df %>% group_by(Species) %>% head()
```

## Summarize

```{r Mean_Length}
# Summarize (mean) using Petal and Sepal Length
iris %>% summarise(Mean.Petal.Length = mean(Petal.Length),
 Mean.Sepal.Length = mean(Sepal.Length))
```

```{r Std.Length}
# Summarize (sd) using Petal and Sepal Length
iris |> summarise(Std.Petal.Length = sd(Petal.Length), Std.Sepal.Length = sd(Sepal.Length))
```

```{r Mean.Width}
# Summarize (mean) using Petal and Sepal Width
iris %>% summarise(Mean.Petal.Width = mean(Petal.Width),
 Mean.Sepal.Width = mean(Sepal.Width))
```

```{r Var.Width}
# Summarize (var) using Petal and Sepal Width
iris %>% summarise(Var.Petal.Width = var(Petal.Width), Var.Sepal.Width = var(Sepal.Width)) 
```

```{r Mean.Species.Petal}
# Summarize (mean) using Petal.Length base on the species column
iris %>% group_by(Species) %>% summarise(Mean.Petal.Length = mean(Petal.Length)) 
```

```{r Var.Species.Petal}
# Summarize (var) using Petal.Length base on the species column
iris %>% group_by(Species) %>% summarise(Var.Petal.Length = var(Petal.Length))
```

```{r Mean.Species.Sepal}
# Summarize (mean) using Sepal.Length base on the species column
iris %>% group_by(Species) %>% summarise(Mean.Sepal.Length = mean(Sepal.Length)) 
```

```{r Median.Species.Petal}
# Summarize (median) using Petal.Width base on the species column
iris |> group_by(Species) |>
 summarize(Median.Petal.Width = median(Petal.Width))
```

```{r Median.Species.Sepal}
# Summarize (median) using Sepal.Width base on the species column
iris %>% group_by(Species) %>% summarize(Median.Sepal.Width = median(Sepal.Width))
```

```{r Total}
# Total number of rows
iris %>% summarise(Observations = n())
```

```{r Each Species}
# Total number of Species base on rows
iris %>% group_by(Species) %>% summarise(Number.Of.Species = n())
```

```{r Samples.Petal}
# Mean of Species base on Petal.Length column
iris %>% group_by(Species) %>% summarise(Number.Of.Samples = n(), Mean.Petal.Length = mean(Petal.Length))
```

```{r Sample.Sepal}
# Mean of Species base on Sepal.Length column
iris %>% group_by(Species) %>% summarise(Number.Of.Samples = n(),
 Mean.Sepal.Length = mean(Sepal.Length))
```

## Breast Cancer Data Manipulation

```{r Class}
format_class <- . %>% {
 dict <- c("2" = "benign", "4" = "malignant")
 dict[as.character(.)]
} %>% unname %>% factor(levels = c("benign", "malignant"))

# Creating Class column from format_classs dict
formatted_breast_cancer <-
 raw_breast_cancer %>% mutate(Class = format_class(Class)) 

# Mean of the cell thickness  base on the class column 
formatted_breast_cancer %>% group_by(Class) %>% summarise(mean.Cl.thickness = mean(Cl.thickness)) 

# Mean of the cell size  base on the class column
formatted_breast_cancer %>% group_by(Class) %>% summarise(mean.Cell.size = mean(Cell.size))
```

```{r Cancer}
# Arrange by Cell.size, group_by Cell.size and Call, summarize by ClassCount
formatted_breast_cancer %>% arrange(Cell.size) %>% group_by(Cell.size, Class) %>% summarise(ClassCount = n()) 
 
formatted_breast_cancer %>% group_by(Class, as.factor(Cell.size)) %>%
 summarise(mean.Cl.thickness = mean(Cl.thickness))

formatted_breast_cancer %>%group_by(as.factor(Cell.size), Class) %>% summarise(mean.Cl.thickness = mean(Cl.thickness)) 
```

```{r Box_plot}
iris %>% select(Species, Petal.Length) %>% head() 
 
# Boxplot with Petal.Length
iris %>% select(Species, Petal.Length) %>% qplot(Species, Petal.Length, geom = "boxplot", data = ., col = Species) 
```

# Tidying Data with tidyr

## Gather

```{r Clean data}
iris %>% gather(key = Attribute, value = Measurement,Sepal.Length, Sepal.Width) %>% select(Species, Attribute, Measurement) %>% head(20) 
```

```{r Plot_facet}
# Boxplot with Sepal length  and  width
iris %>% gather(key = Attribute, value = Measurement, Sepal.Length, Sepal.Width) %>% select(Species, Attribute, Measurement) %>% qplot(Attribute, Measurement, geom = "boxplot", facets = . ~ Species, data = ., col = Species)
```

```{r Clean_2}
iris %>% gather(key = Attribute, value = Measurement,
 Petal.Length, Petal.Width) %>% select(Species, Attribute, Measurement) %>% head(10) 
```

```{r Plot_facet_3}
# Boxplot with Petal length  and  width
iris %>% gather(key = Attribute, value = Measurement, Petal.Length, Petal.Width) %>% select(Species, Attribute, Measurement) %>%
 qplot(Attribute, Measurement, geom = "boxplot",
 facets = . ~ Species, data = ., col = Species)
```

## **Chapter Four**

### Visualizing Data

Visualizing data is a very important part of data science. It helps us
to understand the data better and also helps us to communicate our
findings to others. In this chapter, we will learn how to visualize data
using R.

```{r}
library(ggplot2)
library(graphics)
x = rnorm(100) # random 100 points
y = rnorm(100) # random 100 points
plot(x,y, type ="p", main ="X against Y", col = "lightgreen")
```

```{r}
data(cars) # initiating cars dataset

cars %$% plot(speed, dist, main = "Cars data datasets", xlab = "Speed", ylab = "Stopping distance")

#cars %>% plot(dist ~ speed, data = ., main ="Cars data datasets", col = "red")
```

```{r}
cars %$% plot(speed, dist, main = "Cars data datasets", type = "h",
 xlab = "Speed", ylab = "Stopping Distance") # histogram 

cars %$% plot(speed, dist, main = "Cars data datasets", type = "s",
 xlab = "Speed", ylab = "Stopping Distance") # ste

cars %$% hist(speed, col = "orange", bins = 5) # Histogram of speed

cars %$% hist(dist, col= "blue", bins = 6) # Histogram of distance 

cars %>% lm(dist ~ speed, data = .) %>% plot # Plotting Linear Models
```

```{r}
(data = longley) # initiating longley dataset
#data

longley %>% plot(Unemployed ~ Year, data = ., type = 'l') # plotting the longley dataset
longley %>% lines(Armed.Forces ~ Year, data = ., col = "blue") # adding lines to the plot

longley %$% plot(Unemployed ~ Year, type = 'l',
 ylim = range(c(Unemployed, Armed.Forces))) # plotting the longley dataset
longley %>% lines(Armed.Forces ~ Year, data = ., col = "green") # adding lines to the plot

#as_t(longley)
```

```{r}
cars %>% plot(dist ~ speed, data = ., xlab = "Speed", ylab = "Stopping Distance") # plotting the cars dataset
cars %>% lm(dist ~ speed, data = .) %>% abline(col = "pink") # fitting a line to the plot 
```

```{r}
shape_map <- c("setosa" = 1, "versicolor" = 2,  "virginica" = 3) # encoding the shape of the species

iris %$% plot(Petal.Length ~ Petal.Width, pch = shape_map[Species], col = Species) # plotting the iris dataset

iris %$% plot(Sepal.Length ~ Sepal.Width, pch = shape_map[Species], col = Species) # plotting the iris dataset

#cars %>% qplot(speed, dist, data = .)
```

```{r}
iris %>% qplot(Petal.Width, Petal.Length, color = Species, data = .)  # plotting the iris dataset

iris %>% qplot(Sepal.Width, Sepal.Length, color = Species, data = .)  # plotting the iris dataset

#cars %>% qplot(speed, data = ., bins = 10) # Histogram of speed
```

### Geometries

```{r}

cars %>% qplot(speed, data = ., geom = "density")

cars %>% qplot (dist, data = ., geom = "density")

#cars %>% qplot(speed, dist, data = .)

ggplot(cars) + geom_point(aes(x = speed, y = dist))

#ggplot(cars, aes(x = speed, y = dist)) + geom_point()

#cars %>% ggplot(aes(x = speed, y = dist)) + geom_point()
```

```{r Col.Species}
iris %>% ggplot + geom_point(aes(x = Petal.Width, y = Petal.Length, color = Species))

iris %>% ggplot + geom_point(aes(x = Sepal.Width, y = Sepal.Length,
 color = Species))
```

```{r without.Species}
iris %>% ggplot + geom_point(aes(x = Petal.Width, y = Petal.Length),
 color = "black")

iris %>% ggplot + geom_point(aes(x = Sepal.Width, y = Sepal.Length), col = "brown")
```

```{r Cars}
cars %>% ggplot + geom_histogram(aes(x = speed), bins = 10, col = "lightblue")

cars %>% ggplot + geom_density(aes(x = speed))

cars %>% ggplot + geom_histogram(aes(x=dist), bins = 10, col = "lightpink")

cars %>% ggplot + geom_density(aes(x=dist))
```

```{r}
cars %>% ggplot(aes(x = speed, y = after_stat(count))) + geom_histogram(bins = 10) + geom_density()

cars %>% ggplot(aes(x = dist, y = after_stat(count))) + geom_histogram(bins = 10) + geom_density()
```

```{r Linear models}
cars %>% ggplot(aes(x = speed, y = dist)) + geom_point() + geom_smooth(method = "lm", formula =  y ~ x)

cars %>% ggplot(aes(x = speed, y = dist)) + geom_point() + geom_smooth( method = "loess", formula =  y ~ x) 
```

```{r Longley}
longley %>% ggplot(aes(x = Year)) + geom_line(aes(y = Unemployed), color = "green") + geom_line(aes(y = Armed.Forces), color = "blue") 
```

```{r Line.Point}
longley %>% ggplot(aes(x = Year)) + geom_point(aes(y = Unemployed), color = "green") + geom_point(aes(y = Armed.Forces), color = "blue") + geom_line(aes(y = Unemployed)) + geom_line(aes(y = Armed.Forces), color = "orange") 
```

```{r Legend.1}
longley %>% gather(key, value, Unemployed, Armed.Forces) %>%
 ggplot(aes(x = Year, y = value, color = key)) + geom_line() + geom_point()

longley %>% gather(key, value, Unemployed, Armed.Forces) %>%
 ggplot(aes(x = Year, y = value, color = key)) + geom_line() +
 facet_grid(key ~ .)  # Facet row-wise with lines

#longley %>% gather(key, value, Unemployed, Armed.Forces) %>% gplot(aes(x = Year, y = value, color = key)) + geom_point() + facet_grid(key ~ .) # Facet row-wise with points

```

```{r}
longley %>% gather(key, value, Unemployed, Armed.Forces) %>% ggplot(aes(x = Year, y = value, color = key)) + geom_line() +facet_grid(. ~ key  ) # Facet column-wise with lines

#longley %>% gather(key, value, Unemployed, Armed.Forces) %>% ggplot(aes(x = Year, y = value, color = key)) + geom_point() + facet_grid(. ~ key) # Facet column-wise with points
```

```{r Legend.2}
# row-wise 
iris %>% gather(Measurement, Value, -Species) %>% ggplot(aes(x = Species, y = Value, color = Measurement)) + geom_boxplot() + facet_grid(Measurement ~ .) 

# Column wise
#iris %>% gather(Measurement, Value, -Species) %>% ggplot(aes(x = Species, y = Value,color = Measurement )) + geom_boxplot() + facet_grid(. ~ Measurement ) 
```

```{r Mapped}
iris %>% gather(Measurement, Value, -Species) %>% ggplot(aes(x = Species, y = Value, color = Measurement)) +geom_boxplot() + facet_grid(Measurement ~ ., scale = "free_y") 

label_map <- c(Petal.Width = "Petal Width",Petal.Length = "Petal Length", Sepal.Width = "Sepal Width", Sepal.Length = "Sepal Length")
 
iris %>% gather(Measurement, Value, -Species) %>% ggplot(aes(x = Species, y = Value, color = Measurement)) + geom_boxplot() + facet_grid(Measurement ~ ., scale = "free_y", labeller = labeller(Measurement = label_map))
```

### Scaling

```{r Continuous }
cars %>% ggplot(aes(x = speed, y = dist)) + geom_point() + geom_smooth(method = "lm", formula =  y ~ x) + scale_x_continuous("Speed") + scale_y_continuous("Stopping Distance") 
 
longley %>% gather(key, value, Unemployed, Armed.Forces) %>% ggplot(aes(x = Year, y = value, color = key)) + geom_line() + geom_point() + scale_x_continuous(breaks = 1947:1962) + facet_grid(key ~ .)
```

```{r Reverse}
# Reverse the x axis 
cars |> ggplot(aes(x = speed, y = dist)) + geom_point() + geom_smooth(method = "lm", formula =  y ~ x) + scale_x_reverse("Speed") + scale_y_continuous("Stopping Distance") 

# Reverse the y axis
cars |> ggplot(aes(x = speed, y = dist)) + geom_point() + geom_smooth(method = "lm", formula =  y ~ x) + scale_x_continuous("speed") + scale_y_reverse("Stopping Distance")
```

```{r Species.Modify}
iris |> ggplot(aes(x = Species, y = Petal.Length)) + geom_boxplot() + geom_jitter(width = 0.1, height = 0.1) 

iris |> ggplot(aes(x = Species, y = Petal.Length)) + geom_boxplot() + geom_jitter(width = 0.1, height = 0.1) + scale_x_discrete(labels = c("setosa" = "Setosa", "versicolor" = "Versicolor", "virginica" = "Virginica")) 

iris |> ggplot(aes(x = Species, y = Sepal.Length)) + geom_boxplot() + geom_jitter(width = 0.001, height = 0.001) + scale_x_discrete(labels = c("setosa" = "Setosa", "versicolor" = "Versicolor", "virginica" = "Virginica")) 

#??geom_jitter
```

```{r Color.Species}
iris %>% gather(Measurement, Value, -Species) %>% ggplot(aes(x = Species, y = Value, fill = Species)) + geom_boxplot() + facet_grid(Measurement ~ ., scale = "free_y", labeller = labeller(Measurement = label_map)) + scale_x_discrete(labels = c("setosa" = "Setosa", "versicolor" = "Versicolor", "virginica" = "Virginica")) 

iris %>% gather(Measurement, Value, -Species) %>% ggplot(aes(x = Species, y = Value, fill = Species)) + geom_boxplot() + scale_fill_manual(values = c("orange", "lightgreen", "lightblue")) + facet_grid(Measurement ~ ., scale = "free_y", labeller = labeller(Measurement = label_map)) + scale_x_discrete(labels = c("setosa" = "Setosa", "versicolor" = "Versicolor", "virginica" = "Virginica")) 
```

```{r Greens}
iris %>% gather(Measurement, Value, -Species) %>% ggplot(aes(x = Species, y = Value, fill = Species)) + geom_boxplot() + scale_fill_brewer(palette = "Greens") + facet_grid(Measurement ~ ., scale = "free_y", labeller = labeller(Measurement = label_map)) 

#theme_set(theme_bw(2)) 

#old <- theme_set(theme_bw())
#theme_set(old)
#theme_replace(panel.grid.minor = element_line(colour = "red"))
```

```{r Reds}
iris %>% gather(Measurement, Value, -Species) %>% ggplot(aes(x = Species, y = Value, fill = Species)) + geom_boxplot() + scale_x_discrete(labels = c("setosa" = "Setosa", "versicolor" = "Versicolor", "virginica" = "Virginica")) + scale_fill_brewer(palette = "Reds") + facet_grid(Measurement ~ ., switch = "y", labeller = labeller(Measurement = label_map)) + coord_flip() 
```

```{r Blues}
iris %>% gather(Measurement, Value, -Species) %>%
 ggplot(aes(x = Species, y = Value, fill = Species)) +
 geom_boxplot() + scale_x_discrete(labels = c("setosa" = "Setosa", "versicolor" = "Versicolor", "virginica" = "Virginica")) + scale_fill_brewer(palette = "Blues") +
 facet_grid(Measurement ~ ., switch = "y", labeller = labeller(Measurement = label_map)) + coord_flip()
```

```{r Top.Position}
iris %>% gather(Measurement, Value, -Species) %>%
 ggplot(aes(x = Species, y = Value, fill = Species)) +
 geom_boxplot() + scale_x_discrete(labels = c("setosa" = "Setosa", "versicolor" = "Versicolor", "virginica" = "Virginica")) + scale_fill_brewer(palette = "Greens") +
 facet_grid(Measurement ~ ., switch = "y", labeller = labeller(Measurement = label_map)) + coord_flip() + theme(strip.background = element_blank()) + theme(legend.position ="top")
```

```{r Bottom}
label_map <- c(Petal.Width = "Petal Width",Petal.Length = "Petal Length",Sepal.Width = "Sepal Width",Sepal.Length = "Sepal Length")
 
species_map <- c(setosa = "Setosa",versicolor = "Versicolor",virginica = "Virginica")
 
iris %>% gather(Measurement, Value, -Species) %>%
 ggplot(aes(x = Species, y = Value, fill = Species)) +
 geom_boxplot() + scale_x_discrete(labels = species_map) +
 scale_fill_brewer(palette = "Greens", labels = species_map) + facet_grid(Measurement ~ ., switch = "y",
 labeller = labeller(Measurement = label_map)) +
 coord_flip() + theme(strip.background = element_blank()) +
 theme(legend.position="bottom") 
```

### Figures with Multiple Plots

```{r}
petal <- iris %>% ggplot() + geom_point(aes(x = Petal.Width, y = Petal.Length, color = Species)) + theme(legend.position = "none")
 
sepal <- iris %>% ggplot() + geom_point(aes(x = Sepal.Width, y = Sepal.Length, color = Species)) + theme(legend.position = "none")
 
library(gridExtra)
grid.arrange(petal, sepal, nrow = 2, ncol = 1)

grid.arrange(petal, sepal, nrow = 1, ncol= 2)

library(cowplot)
plot_grid(petal, sepal, labels = c("Petal", "Sepal"))
```

## **Chapter Five**

### Working With Larges Data sets

```{r}
iris %>% sample_n(size = 10) # Randomly sample 10 of iris dataset

iris %>% sample_frac(size = 0.02) # Randomly sample 2% of iris dataset

cars %>% sample_n(size = 25) # Randomly sample 25 of cars dataset

cars %>% sample_frac(size = 0.25) %>% ggplot(aes(x = speed, y = dist)) + geom_point() # Randomly sample 25% of cars dataset

library(pryr)

mem_change(x <- rnorm(10000)) 
mem_change(x[1] <- 0) 
mem_change(y <- x) 
mem_change(x[1] <- 0) 
```

```{r}
d <- data.frame(x = rnorm(10000), y = rnorm(10000))

d %>% ggplot(aes(x = x, y = y)) + geom_point(alpha = 0.2)

mem_change(d)

d %>% ggplot(aes(x = x, y = y)) + geom_density_2d()

mem_change(d)

d |> ggplot(aes(x = x, y = y)) + geom_area()
mem_change(d)
```

```{r}
#install.packages("hexbin")
library(hexbin)
d %>% ggplot(aes(x = x, y = y)) + geom_hex() + scale_fill_gradient(low = "lightgray", high = "red") + geom_density2d(color = "black")

d %>% ggplot(aes(x = x, y = y)) + geom_hex()
```

```{r}
library(ff)

ffcars <- as.ffdf(cars)
summary(ffcars)

ffiris <- as.ffdf(iris)
summary(ffiris)
```

```{r}
library(biglm)

model <- biglm(dist ~ speed, data = ffcars)
summary(model) 

iris_db <- src_sqlite("iris_db.sqlite3", create = TRUE)

iris_sqlite <- tbl(iris_db, "iris", temporary = FALSE)

iris_sqlite %>% group_by(Species) %>% summarise(mean.Petal.Length = mean(Petal.Length),mean.Sepal.Length = mean(Sepal.Length))
```

## **Chapter Six**

## **Supervise Machine Learning**

### Linear Regression

```{r Regression}
library(tidyverse) 
library( magrittr)

cars %>% head()
cars %>% ggplot(aes(x = speed, y = dist)) + geom_point() + geom_smooth(method = "lm", formula = y ~ x) 
cars %>% lm(dist ~ speed, data = .) %>% summary 
cars %>% lm(dist ~ speed, data = .) %>% coefficients 
cars %>% lm(dist ~ speed, data = .) %>% confint 
```

```{r Predict}
predict_dist <- function(speed, theta_1)
 data.frame(speed = speed,
 dist = theta_1 * speed,
 theta = as.factor(theta_1))

cars %>% ggplot(aes(x = speed, y = dist, colour = theta)) +
 geom_point(colour = "black") +
 geom_line(data = predict_dist(cars$speed, 2)) +
 geom_line(data = predict_dist(cars$speed, 3)) +
 geom_line(data = predict_dist(cars$speed, 4)) +
 scale_color_discrete(name = expression(theta[1]))
```

```{r}
thetas <- seq(0, 5, length.out = 50)

fitting_error <- Vectorize(function(theta)
 sum((theta * cars$speed - cars$dist)**2)
) 

data.frame(thetas = thetas, errors = fitting_error(thetas)) %>%
 ggplot(aes(x = thetas, y = errors)) + geom_line() + xlab(expression(theta[1])) + ylab("error") 
```

```{r without y_intercept}
cars %>% lm(dist ~ speed - 1, data = .) %>% coefficients
 
cars %>% ggplot(aes(x = speed, y = dist)) + geom_point() + geom_smooth(method = "lm", formula = y ~ x - 1)
```

### Logistic Regression

```{r jitter}
data("BreastCancer")
BreastCancer %>% head
BreastCancer %>% ggplot(aes(x = Cl.thickness, y = Class, color = Class)) +
 geom_jitter(height = 0.05, width = 0.3, alpha = 0.4)
 
#data("BreastCancer")
#BreastCancer %>% sample_frac(0.4) %>% head
BreastCancer %>% ggplot(aes(x = Cell.size, y = Class, color = Class)) +
geom_jitter(height = 0.05, width = .3, alpha = 0.4)
```

```{r height}
BreastCancer %>% mutate(Cl.thickness.numeric = as.numeric(as.character(Cl.thickness))) %>% mutate(IsMalignant = ifelse(Class == "benign", 0, 1)) %>% ggplot(aes(x = Cl.thickness.numeric, y = IsMalignant)) + geom_jitter(height = 0.05, width = 0.3, alpha = 0.4) + geom_smooth(method = "glm", formula = y ~ x, method.args = list(family = "binomial")) 

BreastCancer %>% mutate(Cl.thickness.numeric =
 as.numeric(as.character(Cl.thickness))) %>%
 mutate(IsMalignant = ifelse(Class == "benign", 0, 1)) %>%
 glm(IsMalignant ~ Cl.thickness.numeric, family = "binomial", data = .) 
```

### Model Matrice and Formula

```{r Model.M}
cars %>% model.matrix(dist ~ speed, data = .) %>% head()

cars %>% model.matrix(dist ~ speed - 1, data = .) %>% head() # without intercept
```

```{r Size + Thickness} }
BreastCancer %>% mutate(Cl.thickness.numeric = as.numeric(as.character(Cl.thickness)), Cell.size.numeric = as.numeric(as.character(Cell.size))) %>% mutate(IsMalignant = ifelse(Class == "benign", 0, 1)) %>% model.matrix(IsMalignant ~ Cl.thickness.numeric + Cell.size.numeric, data = .) %>% head() 
```

```{r without intercept}
BreastCancer %>% mutate(Cl.thickness.numeric = as.numeric(as.character(Cl.thickness)), Cell.size.numeric = as.numeric(as.character(Cell.size))) %>% mutate(IsMalignant = ifelse(Class == "benign", 0, 1)) %>%
 model.matrix(IsMalignant ~ Cl.thickness.numeric + Cell.size.numeric - 1, data = .) %>% head() # without intercept
```

```{r without_inercept}
BreastCancer %>% mutate(Cl.thickness.numeric = as.numeric(as.character(Cl.thickness)), Cell.size.numeric = as.numeric(as.character(Cell.size))) %>% mutate(IsMalignant = ifelse(Class == "benign", 0, 1)) %>%
 glm(IsMalignant ~ Cl.thickness.numeric + Cell.size.numeric - 1, family = "binomial", data = .) # with intercept
```

```{r with_intercept}
 BreastCancer %>% mutate(Cl.thickness.numeric = as.numeric(as.character(Cl.thickness)), Cell.size.numeric = as.numeric(as.character(Cell.size))) %>% mutate(IsMalignant = ifelse(Class == "benign", 0, 1)) %>% glm(IsMalignant ~ Cl.thickness.numeric + Cell.size.numeric, family = "binomial", data = .) # without intercept
```

```{r Bare.nuclei}
BreastCancer %>% mutate(IsMalignant = ifelse(Class == "benign", 0, 1)) %>% model.matrix(IsMalignant ~ Bare.nuclei, data = .) %>% head() 
```

```{r Cl}
BreastCancer %>% mutate(IsMalignant = ifelse(Class == "benign", 0, 1)) %>% model.matrix(IsMalignant ~ Cl.thickness, data = .) %>% head() 
```

```{r thickness*size}
BreastCancer %>% mutate(Cl.thickness.numeric = as.numeric(as.character(Cl.thickness)),Cell.size.numeric = as.numeric(as.character(Cell.size))) %>% mutate(IsMalignant = ifelse(Class == "benign", 0, 1)) %>%
 model.matrix(IsMalignant ~ Cl.thickness.numeric * Cell.size.numeric, data = .) %>% head() 
```

```{r thickness* nuclei}
BreastCancer %>% mutate(Cl.thickness.numeric = as.numeric(as.character(Cl.thickness))) %>% mutate(IsMalignant = ifelse(Class == "benign", 0, 1)) %>% model.matrix(IsMalignant ~ Cl.thickness.numeric * Bare.nuclei, data = .) %>% head() 
```

```{r benign}
BreastCancer %>% mutate(Cl.thickness.numeric = as.numeric(as.character(Cl.thickness))) %>% mutate(IsMalignant = ifelse(Class == "benign", 0, 1)) %>% model.matrix(IsMalignant ~ Cl.thickness.numeric : Bare.nuclei, data = .) %>% head() 
```

```{r poly}
cars %>% model.matrix(dist ~ speed + speed^2, data = .) %>% head 
```

### Polynomial function

```{r poly.1}
cars %>% model.matrix(dist ~ speed + I(speed^2), data = .) %>% head 
 
cars %>% lm(dist ~ speed + I(speed^2), data = .) %>% summary

cars %>% ggplot(aes(x = speed, y = dist)) + geom_point(color = "orange") + geom_smooth(method = "lm", formula = y ~ x + I(x^2))
```

```{r}
iris %>% model.matrix(Sepal.Width ~ Sepal.Length, data = .) %>% head

iris %>% model.matrix(Sepal.Width ~ Sepal.Length + I(Sepal.Length^2), data=.) %>% head()

```

```{r}
iris %>% lm(Sepal.Width ~ Sepal.Length + I(Sepal.Length^2), data = .) %>% summary

iris %>% ggplot(aes(x = Sepal.Width, y = Sepal.Length, col = Species)) + geom_point() + geom_smooth(method = "lm", formula = y ~ x + I(x^2))
```

```{r}
iris %>% model.matrix(Petal.Width ~ Petal.Length + I(Petal.Length**2), data = .) %>% head()

iris %>% ggplot(aes(x = Petal.Width, y = Petal.Length, col = Species)) + geom_point() + geom_smooth(method = "lm", formula = y ~ x + I(x^2))
```

### Validating Models

```{r Mod}
line <- cars %>% lm(dist ~ speed, data = .)

poly <- cars %>% lm(dist ~ speed + I(speed^2), data = .)

predict(line, cars) %>% head

predict(poly, cars) %>% head 
```

### Evaluating Regression Models

```{r rmse}
# checking for root mean square error
rmse <- function(x,t) sqrt(mean(sum((t - x)^2)))

rmse(predict(line, cars), cars$dist)

rmse(predict(poly, cars), cars$dist)
```

```{r train_test}
training_data <- cars[1:25,] # train 25 data set 
test_data <- cars[26:50,] # test 25 data set
line <- training_data %>% lm(dist ~ speed, data = .)
poly <- training_data %>% lm(dist ~ speed + I(speed^2), data = .) 

rmse(predict(line, test_data), test_data$dist) 
rmse(predict(poly, test_data), test_data$dist) 

sampled_cars <- cars %>% mutate(training = sample(0:1, nrow(cars), replace = TRUE)) # 0 = not replace , 1 = replace
sampled_cars %>% head(10) 
```

```{r}
training_data <- sampled_cars %>% filter(training == 1) 
test_data <- sampled_cars %>% filter(training == 0)

training_data %>% head 
test_data %>% head 

line <- training_data %>% lm(dist ~ speed, data = .)
poly <- training_data %>% lm(dist ~ speed + I(speed^2), data = .) 

rmse(predict(line, test_data), test_data$dist) 
rmse(predict(poly, test_data), test_data$dist) 
```

### Evaluating Classification Models

```{r response}
formatted_data <- BreastCancer %>% mutate(Cl.thickness.numeric =
 as.numeric(as.character(Cl.thickness)), Cell.size.numeric =
 as.numeric(as.character(Cell.size))) %>% mutate(IsMalignant = ifelse(Class == "benign", 0, 1))
 
fitted_model <- formatted_data %>% glm(IsMalignant ~ Cl.thickness.numeric + Cell.size.numeric, data = .)

predict(fitted_model, formatted_data, type = "response") %>% head 
```

### Confusion Matrix

```{r }
classify <- function(probability) ifelse(probability < 0.5, 0, 1)
classified_malignant <- classify(predict(fitted_model, formatted_data)) 
table(formatted_data$IsMalignant, classified_malignant) 
```

```{r with_name}
table(formatted_data$IsMalignant, classified_malignant,
 dnn = c("Data", "Predictions"))
```

```{r}
classify <- function(probability) ifelse(probability < 0.5, "benign", "malignant")

classified <- classify(predict(fitted_model, formatted_data))
table(formatted_data$Class, classified, dnn = c("Data", "Predictions")) 
```

### Accuracy

```{r acc}
confusion_ <- table(formatted_data$Class, classified,dnn = c("Data", "Predictions"))

(accuracy <- sum(diag(confusion_))/sum(confusion_))
 
table(BreastCancer$Class) 
```

```{r}
tbl <- table(BreastCancer$Class)
tbl["benign"] / sum(tbl)

tbl = table (BreastCancer$Class)
tbl["malignant"] / sum(tbl)
```

```{r}
table(BreastCancer$Class, sample(BreastCancer$Class)) # random sampling

```

```{r Random_sampling}
accuracy <- function(confusion_matrix)
  sum(diag(confusion_matrix))/sum(confusion_matrix)

replicate(8,accuracy(table(BreastCancer$Class,sample(BreastCancer$Class))))

(accuracy(table(BreastCancer$Class,sample(BreastCancer$Class))))
```

### Sensitivity and Specificity

The `specificity` measure captures how often the model predicts a
negative case correctly and `sensitivity` measure capture how often the
model predicts a positive case correctly.

```{r Neg.Pos}
(specificity <- confusion_[1,1]/
 (confusion_[1,1] + confusion_[1,2])) # true negative class

(sensitivity <- confusion_[2,2]/
 (confusion_[2,1] + confusion_[2,2])) # true positive class
```

```{r random_prediction}
specificity <- function(confusion_) 
  confusion_[1,1]/(confusion_[1,1] + confusion_[1,2])
 
sensitivity <- function(confusion_) 
  confusion_[2,2]/(confusion_[2,1] + confusion_[2,2])
 
prediction_summary <- function(confusion_)
 c("accuracy" = accuracy(confusion_),
 "specificity" = specificity(confusion_),
 "sensitivity" = sensitivity(confusion_))
 
random_prediction_summary <- function()
 prediction_summary(table(BreastCancer$Class,
 sample(BreastCancer$Class)))
 
replicate(3, random_prediction_summary())
```

### Other Measures

```{r Measure}
confusion_[2,1] / sum(confusion_[,1]) # false negative
confusion_[1,1] / sum(confusion_[,1]) # true negative

confusion_[2,2] / sum(confusion_[,2]) # true positive
confusion_[1,2] / sum(confusion_[,2]) # false positive
```

### Random Permutation of Data

```{r Perm}
permuted_cars <- cars[sample(1:nrow(cars)),]
permuted_cars %>% head()

```

```{r}
permute_rows <- function(df) df[sample(1:nrow(df)),]
permuted_cars <- cars %>% permute_rows

group_data <- function(df, n) {
 groups <- rep(1:n, each = nrow(df)/n)
 split(df, groups)
}
cars %>% permute_rows %>% group_data(5) %>% head(1)
```

```{r group_cars, warning=FALSE}
grouped_cars <- cars %>% permute_rows %>% group_data(5)
grouped_cars[[1]] 
grouped_cars[1] 

(grouped_cars[[1]] %>% lm(dist ~ speed, data = .) %>%.$coefficients)
 
#qplot(grouped_cars[[1]] %>%lm(dist ~ speed, data = .) %>%.$coefficients)
```

```{r est}
estimates <- grouped_cars[[1]] %>% lm(dist ~ speed, data = .) %>%.$coefficients

for (i in 2:length(grouped_cars)) {
 group_estimates <- grouped_cars[[i]] %>% lm(dist ~ speed, data = .) %>% .$coefficients
 estimates <- rbind(estimates, group_estimates)
}
estimates 
```

```{r esti}
library(purrr) 

estimates <- grouped_cars %>% map(. %>% lm(dist ~ speed, data = .) %>% .$coefficients)
 
estimates 

estimates <- grouped_cars %>% map(. %>% lm(dist ~ speed, data = .) %>% .$coefficients) %>% do.call("rbind", .)
 
estimates
```

### Cross Validation

```{r valid}
cross_validation_groups <- function(grouped_df) {
 result <- vector(mode = "list", length = length(grouped_df))
 for (i in seq_along(grouped_df)) {
 result[[i]] <- grouped_df[-i] %>% do.call("rbind", .)
 }
 result
} 

cars %>% permute_rows %>% group_data(5) %>% cross_validation_groups   %>% map(. %>% lm(dist ~ speed, data = .) %>% .$coefficients) %>% do.call("rbind", .)
```

```{r split}
cross_validation_split <- function(grouped_df) {
 result <- vector(mode = "list", length = length(grouped_df))
 for (i in seq_along(grouped_df)) {
 training <- grouped_df[-i] %>% do.call("rbind", .)
 test <- grouped_df[[i]]
 result[[i]] <- list(training = training, test = test)
 }
 result
}

cars %>% permute_rows %>% group_data(2) %>% cross_validation_split
```

```{r}
prediction_accuracy_cars <- function(test_and_training) {
 result <- vector(mode = "numeric",
 length = length(test_and_training))
 for (i in seq_along(test_and_training)) {
 training <- test_and_training[[i]]$training
 test <- test_and_training[[i]]$test
 model <- training %>% lm(dist ~ speed, data = .)
 predictions <- test %>% predict(model, data = .)
 targets <- test$dist
 result[i] <- rmse(targets, predictions)
 }
 result
}

cars %>% permute_rows %>%group_data(2) %>% cross_validation_split %>% prediction_accuracy_cars
```

### Selecting Random Training and Testing Data

```{r}
random_group <- function(n, probs) {
 probs <- probs / sum(probs)
 g <- findInterval(seq(0, 1, length = n), c(0, cumsum(probs)),
 rightmost.closed = TRUE)
 names(probs)[sample(g)]
}

random_group(8, c(training = 0.5, test = 0.5)) # 50/50

random_group(8, c(training = 0.6, test = 0.4)) # 60/40

random_group(8, c(training = 0.8, test = 0.2)) # 80/20

```

```{r}
partition <- function(df, n, probs) {
 replicate(n, split(df, random_group(nrow(df), probs)), FALSE)
}

random_cars <- cars %>% partition(4, c(training = 0.5, test = 0.5))

random_cars %>% prediction_accuracy_cars 
```

### Examples of Supervised Learning Packages

```{r Tree}
### Decision Tree
library(rpart)

# regression
model <- cars %>% rpart(dist ~ speed, data = .)
rmse(predict(model, cars), cars$dist) # root mean square

# classification
model <- BreastCancer %>%rpart(Class ~ Cl.thickness, data = .) 
predict(model, BreastCancer) %>% head 

predicted_class <- predict(model, BreastCancer) %>%
 as.data.frame %$% ifelse(benign > 0.5, "benign", "malignant")
 
table(BreastCancer$Class, predicted_class)

accuracy = sum(453,147) / sum(453,147,5,94)
accuracy
```

```{r}
# regression
model <- rock %>% rpart(shape ~ perm, data = .)
rmse(predict(model, rock), rock$perm) # root mean square

# regression
model <- women %>%rpart(height ~ weight, data = .) 
rmse(predict(model, women), women$weight)

#require(graphics)
pairs(quakes, main = "Fiji Earthquakes, N = 1000", cex.main = 1.2, pch = ".")

model <- quakes %>%rpart(depth ~ mag, data = .) 
rmse(predict(model, quakes), quakes$mag)
```

```{r Part}
#install.packages("party")
library(party)
# regression
model <- cars %>% ctree(dist ~ speed, data = .)
rmse(predict(model, cars), cars$dist) 

# classification
model <- BreastCancer %>% ctree(Class ~ Cl.thickness, data = .) 

predict(model, BreastCancer) %>% head 

table(BreastCancer$Class, predict(model, BreastCancer)) 

cars %>% ctree(dist ~ speed, data = .) %>% plot
```

```{r Forest}
# Random Forests
library(randomForest)
# regression
model <- cars %>% randomForest(dist ~ speed, data = .)
rmse(predict(model, cars), cars$dist) # root mean square error
 
# classification
model <- BreastCancer %>% randomForest(Class ~ Cl.thickness, data = .)
 
predict(model, BreastCancer) %>% head() 

table(BreastCancer$Class, predict(model, BreastCancer))

accuracy = sum(437,165) / sum(437,165,76,21)
accuracy
```

```{r Neural}
# Neural Networks
library(nnet)
model <- cars %>% nnet(dist ~ speed, data = ., size = 5)
rmse(predict(model, cars), cars$dist)

model <- BreastCancer %>% nnet(Class ~ Cl.thickness, data = ., size = 5)

predict(model, BreastCancer) %>% head 

predicted_class <- predict(model, BreastCancer) %>%
 { ifelse(. < 0.5, "benign", "malignant") }
 
table(BreastCancer$Class, predicted_class)

accuracy = sum(437,165) / sum(437,165,76,21)
accuracy
```

```{r Machines}
# Support Vector Machines
#install.packages("kernlab")
library(kernlab)

# regression
model <- cars %>% ksvm(dist ~ speed, data = .)
rmse(predict(model, cars), cars$dist)

# classification
model <- BreastCancer %>% ksvm(Class ~ Cl.thickness, data = .)

predict(model, BreastCancer) %>% head 

table(BreastCancer$Class, predict(model, BreastCancer))

(accuracy = sum(437,165)/ sum(437,165,21,76))
```

```{r Naive}
### Naive Bayes
library(e1071)

# classification
model <- BreastCancer %>% naiveBayes(Class ~ Cl.thickness, data = .)

predict(model, BreastCancer) %>% head

table(BreastCancer$Class, predict(model, BreastCancer))
```

### **Unsupervised Learning**

```{r Pca}
## **Chapter Seven** 
require(graphics)
### Principal Component Analysis
iris %>% head()

#iris %>% ggplot() + geom_point(aes(x = Sepal.Length, y = Sepal.Width, colour = Species))

#iris %>% ggplot() + geom_point(aes(x = Petal.Length, y = Petal.Width, colour = Species))

pca <- iris %>% select(-Species) %>% prcomp
pca
pca %>% plot

#?prcomp
```

```{r Pred}
mapped_iris <- pca %>% predict(iris)
mapped_iris %>% head

mapped_iris %>% as.data.frame %>% cbind(Species = iris$Species) %>%
ggplot() + geom_point(aes(x = PC1, y = PC2, colour = Species))
```

```{r house}
data(HouseVotes84)
HouseVotes84 %>% head(na.rm = TRUE)

vote_patterns <- HouseVotes84 %>% select(-Class) %>% apply(c(1,2), . %>% { ifelse(as.character(.) == "n", 0, 1) }) %>% apply(c(1,2), . %>% { ifelse(is.na(.), 0.5, .) })

pca <- vote_patterns %>% prcomp
#pca %>% head
pca %>% plot
```

```{r votes}
mapped_votes <- pca %>% predict(vote_patterns)

mapped_votes %>%as.data.frame %>% cbind(Class = HouseVotes84$Class) %>% ggplot() + geom_point(aes(x = PC1, y = PC2, colour = Class))
```

```{r dimen}
### Multidimensional Scaling 
iris_dist <- iris %>% select(-Species) %>% dist

mds_iris <- iris_dist %>% cmdscale(k = 2)

mds_iris %>% head

mds_iris %>% as.data.frame %>% cbind(Species = iris$Species) %>% ggplot() + geom_point(aes(x = V1, y = V2, colour = Species))
```

```{r mds}
mds_votes <- vote_patterns %>% dist %>% cmdscale(k = 2 )

mds_votes %>% as.data.frame %>% cbind(Class = HouseVotes84$Class) %>% ggplot() + geom_point(aes(x = V1, y = V2, colour = Class))
```

```{r}
random_ngram <- function(n)
 sample(c('C','H','A','R','L','E','S'), size = n, replace = TRUE) %>% paste0(collapse = "")

random_string <- function(m) {
 n <- max(1, m + sample(c(-1,1), size = 1) * rgeom(1, 1/2))
 random_ngram(n)
}

strings <- replicate(10, random_string(5))
#install.packages("stringdist")
library(stringdist)
string_dist <- stringdistmatrix(strings)

string_dist %>% cmdscale(k = 4) %>% as.data.frame %>%
 cbind(String = strings) %>% ggplot(aes(x = V1, y = V2)) +
 geom_point() + geom_label(aes(label = String), hjust = 0, nudge_y = -0.1)
```

### Clustering

Clustering methods seek to find similarities between data points and
group data according to these similarities.

```{r clust}
### Kmeans
clusters <- iris %>% select(-Species) %>% kmeans(centers = 3)

clusters$centers

clusters$cluster %>% head

clusters$cluster %>% table

iris %>% cbind(Cluster = clusters$cluster) %>%
 ggplot() + geom_bar(aes(x = Species, fill = as.factor(Cluster)), position = "dodge") + scale_fill_discrete("Cluster")
```

```{r cent}
pca <- iris %>% select(-Species) %>% prcomp
 
mapped_iris <- pca %>% predict(iris)
 
mapped_centers <- pca %>% predict(clusters$centers)

mapped_iris %>% as.data.frame %>% cbind(Species = iris$Species,
 Clusters = as.factor(clusters$cluster)) %>% ggplot() + geom_point(aes(x = PC1, y = PC2, colour = Species, shape = Clusters)) + geom_point(aes(x = PC1, y = PC2), size = 5, shape = "X", data = as.data.frame(mapped_centers))

table(iris$Species, clusters$cluster)

tbl <- table(iris$Species, clusters$cluster)
(counts <- apply(tbl, 1, which.max))

tbl <- table(iris$Species, clusters$cluster)
(counts <- apply(tbl, 1, which.min))

map <- rep(NA, each = 3)
map[counts] <- names(counts)
table(iris$Species, map[clusters$cluster])

#?prcomp
```

```{r hier}
## Hierarchical Clustering
iris_dist <- iris %>% select(-Species) %>% scale %>% dist

clustering <- hclust(iris_dist)

plot(clustering)
```

```{r gend}
library(ggdendro)
ggdendrogram(clustering) + theme_dendro()

clusters <- clustering %>% cutree(k = 3)

iris %>% cbind(Cluster = clusters) %>% ggplot() + geom_bar(aes(x = Species, fill = as.factor(Cluster)), position = "dodge") + scale_fill_discrete("Cluster")
```

```{r map_iris}
mapped_iris %>% as.data.frame %>% cbind(Species = iris$Species,
 Clusters = as.factor(clusters)) %>% ggplot() + geom_point(aes(x = PC1, y = PC2, shape = Species, colour = Clusters))
```

```{r message=FALSE, warning=FALSE}
## Association Rules
library(arules)

data(income)
income %>% head

data(Income)
Income %>% head

rules <- income %>% apriori
rules %>% head %>% inspect(linebreak = FALSE)

rules %>% sort(by = "lift") %>% head %>% inspect(linebreak = FALSE)

rules %>% subset(support > 0.5) %>% sort(by = "lift") %>% head %>% inspect(linebreak = FALSE)
```
# **Chapter 8**
## **More on R Programming**
```{r More}

3 %% 4

3 / 4

3/4

1 : 2 ^ 4

-1 : 2
```

```{r logic}
!TRUE

!FALSE

x <- c(TRUE, FALSE, TRUE, FALSE)

y <- c(TRUE, TRUE, FALSE, FALSE)

x | y

FALSE | TRUE

FALSE || TRUE
```

```{r num}
is.numeric(2)

x <- as.integer(2)

is.integer(x)

as.integer(3.2)

as.character(9.99999)

sqrt(as.complex(1))

(x <- 5 > 4)

class(x)
is.logical(x)
```

```{r hello}
x <- "hello, world"
class(x)
is.character(x)

as.character(3.14) 

rep("foo", 10) # repeat foo 10x

v <- 1:3
is.atomic(v)

v <- 1:3
is.vector(v)
attr(v, "foo") <- "bar"
v
is.vector(v)
```

```{r concat}
c(1, 2, c(3, 4), c(5, 6, 7))

(c(1, 2, 3, "foo") -> vec)

is.vector(vec)
is.character(vec)
```

```{r att}
v <- 1:6

attributes(v)

dim(v) <- c(2, 3)

attributes(v)

dim(v)
matrix(v,nrow = 2, ncol = 3)
```

```{r mat}
v <- 1:6
matrix(v, nrow = 2, ncol = 3, byrow = FALSE)
matrix(v, nrow = 2, ncol = 3, byrow = TRUE)

(A <- c(1:4, nrow = 2))
(B <- c(5:8, nrow = 2))

A * B # element-wise mult
A %*% B # matrix mult

t(A)
# solve.default(A) ### inverse
# solve(A) %*% A ### identity 
```

```{r}
list(1:3, 5:8)

list(1:3, c(TRUE, FALSE))

list((1:4), (5:8), (9:12))

unlist(list(1:4, 5:7))
#?'%%'
# ?`[[`
```

```{r index}
v <- 1:4
v[2]
v[2:3]
v[c(1,1,4,3,2)]
v[-1]
v[-(1:2)]
v[v %% 2 == 0]
v[v %% 2 == 0] <- 13
v
```

```{r drop}
m <- matrix(1:6, nrow = 2, byrow = TRUE)
m
m[1,]
m[,1]
m[1:2,1:2]
m[1,,drop = FALSE]
m[,1,drop = FALSE]
```

```{r}
L <- list(1,2,3)
L[1]
L[2:3]
L[[1]]
L[[2]]
L[[3]]
```
```{r Named}
### Named Values
v <- c(a = 1, b = 2, c = 3, d = 4)
v
L <- list(a = 1:5, b = c(TRUE, FALSE))
L
names(v) <- LETTERS[1:4]
v
v["A"]
L["a"]
L[["a"]] # select only value a
L$a # same
```

```{r control}
### Control Structures
if (10 < 12) {
 print("Yes")
} else {
 print("No")
}

x = 100
if (x < 0) {
 print("Negative")
} else if (x > 0) {
 print("Positive")
} else {
 print("Nothing")
}
```

```{r structure}
if (x > 0) "positive" else if (x < 0) "negative" else "zero"

if (x > 0) {
 print("positive")
}else if (x < 0){
 print("negative")
}else{
 print("zero")
}

for (i in 1:10) {
 print(i * i)
}
```

```{r foo_bar_baz}
x <- c("foo", "bar", "baz")
for (i in seq_along(x)) {
 print(i)
 print(x[i])
}

for (i in seq_along(x)) {
 if (i %% 2 == 0) {
 next
 }
 print(x[i])
}
```

```{r divided_two}
for (i in 1:100) {
 if (i %% 2 == 0) {
 next
 }
 if (i > 20) {
 break
 }
 print(i)
}
```

```{r while_loop}
i <- 1
while (i < 5) {
 i <- i + 1
 print(i)
}

i <- 1
repeat {
 print(i)
 i <- i + 1
 if (i > 5) break
}
```

```{r world}
### Function
f <- function() {
 print("hello, world")
 ((5 + 4) + (4 ** 2)) / 5
}
f()

# 25 / 5
```

```{r Plus_div}
plus <- function(x, y) {
 print(paste(x, "+", y, "is", x + y))
 x + y
}
plus(2, 2)

div <- function(x, y) {
 print(paste(x, "/", y, "is", x / y))
 x / y}

div (6,2)
div(3,4)
div(y = 2, x = 6)
```

```{r}
pow <- function(x, y = 2) x^y
pow(2)
pow(3)
pow(4)
pow(8)
pow(9,2)
pow(8,0)

```
```{r safe}
safer_div <- function(x, y) {
 if (y == 0) {
 NA
 } else {
 x / y
 }
}
safer_div(2,4)
safer_div(3,9)

safer_div <- function(x, y) {
 if (y == 0) {
 return(NA)
 }
 x / y
}
safer_div(3,6)
safer_div(1,0)
```

```{r func}
f <- function(x, y = x^2) y + x
f(7)

g <- function(x, y = x^2) { x <- 0; y + x }
g(2)
g(2,9)

h <- function(x, y = x^2) { y; x <- 0; y + x }
h(2)
h(8)
```
### Scoping
Scope in R is lexical. This means that if a variable is used in a function but not defined in the function or 
part of the functions parameters, then R will start searching outward in the code from where the function 
was created. This essentially means searching outward and upward from the point in the code where the 
function is specified, since a function is created when the code is executed where the function is defined.
```{r scop_1}
x <- "x"
f <- function(y) {
 g <- function() c(x, y)
 g()
}
f("y")

x <- "x"
f <- function(y) {
 g <- function() c(x, y)
 y <- "z"
 g()
}
f("y")
```

```{r scop_2}
x <- "x"
f <- function(y) {
 g <- function() c(x, y)
 g
}
g <- f("y")
g()

x <- "x"
f <- function(y) {
 g <- function() c(x, y)
 g
}
g <- f("u")
g()

h <- f("z")
h()
```

```{r null}
f <- function() {
 x <- NULL
 set <- function(val) { x <- val }
 get <- function() x
 list(set = set, get = get)
}
x <- f()
x$get()

x$set(5)
x$get()
```

```{r}
f <- function(){
x <- NULL
 set <- function(val) { x <<- val }
 get <- function() x
 list(set = set, get = get)
}
x <- f()
x$get()

x$set(5)
x$get()
```

```{r fact}
n <- function(x) x
f <- function(n) n(n)
f(5)

f(function(x) 15)

factorial <- function(n) {
 if (n == 1) {
 1
 } else {
 n * factorial(n - 1)
 }
}
factorial(9)
factorial(1)
```

```{r merge}
merge <- function(x, y) {
 if (length(x) == 0) return(y)
 if (length(y) == 0) return(x)
 if (x[1] < y[1]) {
 c(x[1], merge(x[-1], y))
 } else {
 c(y[1], merge(x, y[-1]))
 }
}
merge(9,0)
merge(12,5)
merge(23,9)
```

```{r merge_sort}
merge_sort <- function(x) {
 if (length(x) < 2) return(x)
 n <- length(x)
 m <- n %/% 2
 merge(merge_sort(x[1:m]), merge_sort(x[(m+1):n]))}
merge_sort (10)

# 3 %% 4

# 3 / 4

# ?'%%'
```

# **Chapter 9**
### **Advance R Programming**
```{r prog}
(x <- 2 / 3)

(y <- x ** 2)

(x <- 1:4 / 3)

(y <- x ** 2)
```

```{r vect}
x <- 1:5
y <- 6:10
(z <- x + y)
(s = y -  x)

z <- vector(length = length(x))
for (i in seq_along(x)) {
 z[i] <- x[i] + y[i]
}
z

s = vector(length = length(x))
for (i in seq_along(x)){
  s[i] = y[i] - x[i]
}
s
```

```{r pow_}
sqrt((1:5)**2)

sin(sqrt((1:5)**2))

x <- 1:10
y <- 1:2
x + y
```

```{r if_else}
x <- 1:10
ifelse(x %% 2 == 0, 5, 15)

f <- function(x, y) sqrt(x ** y)
f(1:6, 2)

f(1:6, 1:2)
```

```{r role}
role_table <- list("Thomas" = "Instructor",
 "Henrik" = "Student",
 "Kristian" = "Student",
 "Randi" = "Student",
 "Heidi" = "Student",
 "Manfred" = "Student",
 "Charles" = "Researcher")

map_to_role <- function(name) role_table[[name]]

# map_to_role("Thomas")

# map_to_role("Charles")

role_table[c("Thomas", "Henrik", "Randi","Charles")]

map_to_role_2 <- function(names) unlist(role_table[names])
x <- c("Thomas", "Henrik", "Randi","Charles")
map_to_role_2(x)

```

```{r lis}
x <- list("first" = list("second" = "foo"), "third" = "bar")
x[[c("first", "second")]]

#x[["first"]][["second"]]
```

```{r apply}
### apply
m <- matrix(1:6, nrow = 2, byrow = TRUE)
m

apply(m, 1, function(x) paste(x, collapse = ":")) # row-wise

apply(m, 2, function(x) paste(x, collapse = ":")) # column-wise

apply(m, c(1,2), function(x) paste(x, collapse = ":")) # Both

apply(m, 1, function(x) c(x,x))

apply(m, 2, function(x) c(x,x))

apply(m, c(1,2), function(x) c(x,x))
```

```{r apply_p}
x <- apply(m, c(1,2), function(x) c(x,x))
k <- dim(x)[3]
n <- dim(x)[2]
for (i in 1:n) {
 for (j in 1:k) {
 print(x[,i,j])
 }
}
```

```{r sumpow}
sumpow <- function(x, n) sum(x) ** n
apply(m, 1, sumpow,n = 2) # row-wise
apply(m, 2, sumpow, n = 2) # column-wise
```

```{r}
cubepow <- function(x, n) cube(x) ** n
apply(m, 1, cubepow, n = 3)
apply(m, 2,cubepow, n = 3)
```
```{r lapply}
### lapply
l <- list(1, 2, 3)
lapply(l, function(x) x**2)

l <- list(a = 1, b = 2, c = 3)
lapply(l, function(x) x**2)

lapply(1:5, function(x) x**4)

lapply(list(a = 1:3, b = 4:6), function(x) x**2)

lapply(list( cv = 1:8), function(x) x**3)
```

```{r sapply}
### sapply
sapply(1:3, function(x) x**2) # returns vector by default 

vapply(1:3, function(x) x**2,1)

#?vapply
```

```{r adv}
### Advance Functions

`+`(2, 2)
v <- 1:4
names(v) <- c("a", "b", "c", "d")
v

`foo<-` <- function(x, value) {
 x$foo <- value
 x
}

`bar<-` <- function(x, value) {
 x$bar <- value
 x
}

```

```{r}
x <- list(foo = 1, bar = 2)
x$foo 
x$bar
foo(x) <- 4
x$foo
bar(x) <- 3
x$bar
```
```{r foo_bar}
y <- x
foo(x) <- 5
x
y

`foo<-` <- function(y, value) {
 y$foo <- value
 y
}

x <- list(foo = 1, bar = 2)
foo(x) <- 3
x$foo
```

```{r foo_bar_1}
`modify<-` <- function(x, variable, value) {
 x[variable] <- value
 x
}

x <- list(foo = 1, bar = 2)
modify(x, "foo") <- 5
modify(x, "bar") <- 7
x$foo
x$bar
```

```{r index}
x <- 1:4
f <- function(x) {
 x[2] <- 5 # insert 5 at index 2
 x[1] <- 9 # insert 9 at index 1 
 x[4] <- 10 # insert 10 at index 4
 x
}
x
f(x)
```

```{r square}
square <- function(x) (x**2 - x)
square(5)

m <- matrix(1:6, nrow = 3)
m

sum_of_squares <- function(x) sum(x^2)
apply(m, 1, sum_of_squares) # row - wise
apply(m,2 , sum_of_squares) # column - wise

#apply(m^2, 1, sum)
```

```{r apply_if}
apply_if <- function(x, p, f) {
 result <- vector(length = length(x))
 n <- 0
 for (i in seq_along(x)) {
 if (p(x[i])) {
 n <- n + 1
 result[n] <- f(x[i])
 }
 }
 head(result, n)
}
apply_if(1:8, function(x) x %% 2 == 0, function(x) x^2)
```

```{r power_cube}
power <- function(n) function(x) x**n
square <- power(2)
cube <- power(3)

x <- 1:4
square(x) 
cube(x)
```

```{r filter_map_reduce}
### Filter, Map and Reduce
is_even <- function(x) x %% 2 == 0
Filter(is_even, 1:8)
Filter(is_even, unlist(1:8))
Filter(is_even, as.list(1:8))

```

```{r}
divi <- function(x) (x %% 3 == 0)
Filter(divi, 1:12)

square <- function(x) x^2
Map(square, 1:4)

unlist(Map(square, 1:4))
```

```{r map}
plus <- function(x, y) x + y
unlist(Map(plus, 0:3, 3:0))

add_parenthesis <- function(a, b) paste("(", a, ", ", b, ")", sep = "")
Reduce(add_parenthesis, 1:4)

mysum <- function(x) Reduce(`+`, x)
mysum(1:4)
```

```{r factorail}
### Function Operations
cached <- function(f) {
 force(f)
 table <- list()
 
function(n) {
 key <- as.character(n)
 if (key %in% names(table)) {
 print(paste("I have already computed the value for", n))
 table[[key]]
 } else {
 print(paste("I'm going to compute the value for", n))
 res <- f(n)
 print(paste("That turned out to be", res))
 table[key] <<- res
 print(table)
 res
 }
 }
}

factorial <- function(n) {
 if (n == 1) {
 1
 } else {
 n * factorial(n - 1)
 }
}

factorial <- cached(factorial)

factorial(4)

# factorial(1)

# factorial(2)

# factorial(3)

# factorial(4)

factorial(5)
```

```{r fibonacci}
fibonacci <- function(n) {
 if (n == 1 || n == 2) {
 1
 } else {
 fibonacci(n - 1) + fibonacci(n - 2)
 }
}

fibonacci <- cached(fibonacci)

fibonacci(4)

fibonacci(1)

fibonacci(2)

fibonacci(3)

fibonacci(4)

fibonacci(5)

# ??fibonacci
```

```{r ellip}
### Ellipsis Parameters
g <- function(a, b, ...) NULL
g(a = 1, b = 2, c = 3)

tolist <- function(...) list(...)
tolist()

# tolist(a = 1)

# tolist(a = 1, b = 3)
tolist(a = 1, b = 3, c = 4,d = 12, e = 23)

```

```{r}
### System Timeout
time_out <- function(f) {
 force(f)
 function(...) {
 system.time(f(...))
 }
}

ti_mean <- time_out(mean)
ti_mean(runif(1e6))

```

```{r blm}
# **Chapter 10**
## **Object Oriented Programming**
### **Bayesian Linear Models**
blm <- function(model, alpha = 1, beta = 1, ...) {
 # Here goes the mathematics for computing the fit.
 frame <- model.frame(model, ...)
 phi <- model.(frame)
 no_params <- ncol(phi)
 target <- model.response(frame)
 covar <- solve(diag(alpha, no_params) + beta * t(phi) %*% phi)
 mean <- beta * covar %*% t(phi) %*% target
 list(formula = model,
 frame = frame,
 mean = mean,
 covar = covar)
}
```

```{r fake}
# fake some data for our linear model
x <- rnorm(10)
a <- 1 ; b <- 1.3
w0 <- 0.2 ; w1 <- 3
y <- rnorm(10, mean = w0 + w1 * x, sd = sqrt(1/b))
#frame <- data.frame (x_1 = x, y_1 = y )
#frame

# fit a model
model <- blm(y ~ x, alpha = a, beta = b)

model

class(model)

class(model) <- "blm"

class(model)
```

```{r simple}
blm <- function(model, alpha = 1, beta = 1, ...) {
 # stuff happens here...
 object <- list(formula = model,
 frame = frame,
 mean = mean,
 covar = covar)
 class(object) <- "blm"
 object
}

blm <- function(model, alpha = 1, beta = 1, ...) {
 # stuff happens here...
 structure(list(formula = model,
 frame = frame,
 mean = mean,
 covar = covar),
 class = "blm")
}

model
```

```{r morphic Function}
### Polymorphic Function
print.blm <- function(x, ...) {
 print(x$formula)
}
model
#??print
```

```{r use_method}
### Defining your own Function
Charles <- function(x, ...) UseMethod("Charles")
Charles.default <- function(x, ...) print("default Charles")
Charles("a string")
#Charles(12)

Charles.blm <- function(x, ...) print("blm Charles")
Charles(model)
```

```{r foo.blm}
Charles.blm <- function(x, upper = FALSE, ...) {
 if (upper) {
 print("BLM FOO")
 } else {
 print("blm foo")
 }
}

Charles("a string")

Charles(model)

Charles("a string", upper = TRUE)

Charles(model, upper = TRUE)
```

```{r foo}
foo <- function(object, ...) UseMethod("foo")
foo.default <- function(object, ...) stop("foo not implemented")

bar <- function(object, ...) UseMethod("bar")
bar.default <- function(object, ...) stop("bar not implemented")


A <- function(f, b) structure(list(foo = f, bar = b), class ="A")

foo.A <- function(object, ...) paste("A::foo ->", object$foo)
bar.A <- function(object, ...) paste("A::bar ->", object$bar)

a <- A("Charles", "Nana")
foo(a)

bar(a)

baz <- function(object, ...) UseMethod("baz")
baz.default <- function(object, ...) stop("baz not implemented")
```

```{r bar.baz}
B <- function(f, b, bb) {
 a <- A(f, b)
 a$baz <- bb
 class(a) <- "B"
 a
}

bar.B <- function(object, ...) paste("B::bar ->", object$bar)
baz.B <- function(object, ...) paste("B::baz ->", object$baz)

B <- function(f, b, bb) {
 a <- A(f, b)
 a$baz <- bb
 class(a) <- c("B", "A")
 a
}

b <- B("Charles", "Kwame", "Appiah")

foo(b)

bar(b)

baz(b)
```

```{r foo.bar.baz}
C <- function(f, b, bb) {
 b <- B(f, b, bb)
 class(b) <- c("C", "B", "A")
 b
}
c <- C("foo", "bar", "baz")
foo(c)

bar(c)

baz(c)

C <- function(f, b, bb) {
 b <- B(f, b, bb)
 class(b) <- c("C", class(b))
 b
}
```

```{r}
# **Chapter 12**
## Testing R packages
area <- function(x) UseMethod("area")

circumference <- function(x) UseMethod("circumference")

rectangle <- function(width, height) {
 structure(list(width = width, height = height),
 class = c("rectangle", "shape"))
}

area.rectangle <- function(x) x$height * x$width

circumference.rectangle <- function(x) 2 * x$height + 2 * x$width

r <- rectangle(width = 3, height = 4)
area(r)

circumference(r)

# Automating testing
r <- rectangle(width = 3, height = 4)
if (area(r) != 3*4) {
 stop("Area not computed correctly!")
}

if (circumference(r) != 2*3 + 2*4) {
 stop("Circumference not computed correctly!")
}
```

```{r}
## **Chapter 14**
### Profiling and  Optimizing
library(profvis)
graph <- function(n, edges) {
 m <- (0, nrow = n, ncol = n)
 no_edges <- length(edges)
 if (no_edges >= 1) {
 for (i in seq(1, no_edges, by = 2)) {
 m[edges[i], edges[i+1]] <- m[edges[i+1], edges[i]] <- 1
 }
 }
 structure(m, class = "graph")
}
```

```{r}
smooth_weights <- function(graph, node_weights, alpha) {
 if (length(node_weights) != nrow(graph))
 stop("Incorrect number of nodes")
 no_nodes <- length(node_weights)
 new_weights <- vector("numeric", no_nodes)
 for (i in 1:no_nodes) {
 neighbour_weights <- 0
 n <- 0
 for (j in 1:no_nodes) {
 if (i != j && graph[i, j] == 1) {
 neighbour_weights <- neighbour_weights + node_weights[j]
 n <- n + 1
 }
 }
 if (n > 0) {
 new_weights[i] <-
 alpha * node_weights[i] +
 (1 - alpha) * neighbour_weights / n
 } else {
 new_weights[i] <- node_weights[i]
 }
 }
 new_weights
}
```

```{r message=FALSE, warning=FALSE}
profvis::profvis({
 n <- 1000
 nodes <- 1:n
 edges <- sample(nodes, 600, replace = TRUE)
 weights <- rnorm(n)
 g <- graph(n, edges)
 smooth_weights(g, weights, 0.8)
})
```

```{r}
graph <- function(n, edges) {
 neighbours <- vector("list", length = n)
 for (i in seq_along(neighbours)) {
 neighbours[[i]] <- vector("integer", length = 0)
 }
 no_edges <- length(edges)
 if (no_edges >= 1) {
 for (i in seq(1, no_edges, by = 2)) {
 n1 <- edges[i]
 n2 <- edges[i+1]
 neighbours[[n1]] <- c(n2, neighbours[[n1]])
 neighbours[[n2]] <- c(n1, neighbours[[n2]])
 }
 }
 for (i in seq_along(neighbours)) {
 neighbours[[i]] <- unique(neighbours[[i]])
 }
 structure(neighbours, class = "graph")
}
```

```{r}
smooth_weights <- function(graph, node_weights, alpha) {
 if (length(node_weights) != length(graph))
 stop("Incorrect number of nodes")
 no_nodes <- length(node_weights)
 new_weights <- vector("numeric", no_nodes)
 for (i in 1:no_nodes) {
 neighbour_weights <- 0
 n <- 0
 for (j in graph[[i]]) {
 if (i != j) {
   neighbour_weights <- neighbour_weights + node_weights[j]
 n <- n + 1
 }
 }
 if (n > 0) {
 new_weights[i] <-
 alpha * node_weights[i] +
 (1 - alpha) * neighbour_weights / n
 } else {
 new_weights[i] <- node_weights[i]
 }
 }
 new_weights
}
```

```{r}
profvis::profvis({
 n <- 10000
 nodes <- 1:n
 edges <- sample(nodes, 600, replace = TRUE)
 weights <- rnorm(n)
 g <- graph(n, edges)
 smooth_weights(g, weights, 0.8)
})
```

```{r}
graph <- function(n, edges) {
 neighbours <- vector("list", length = n)
 for (i in seq_along(neighbours)) {
 neighbours[[i]] <- vector("integer", length = 0)
 }
 no_edges <- length(edges)
 if (no_edges >= 1) {
 sources <- seq(1, no_edges, by = 2)
 destinations <- seq(2, no_edges, by = 2)
 edge_ <- (NA, nrow = length(sources), ncol = 2)
 edge_[,1] <- edges[sources]
 edge_[,2] <- edges[destinations]
 for (i in 1:nrow(edge_)) {
 if (edge_[i,1] > edge_[i,2]) {
 edge_[i,] <- c(edge_[i,2], edge_[i,1])
 }
 }
 edge_ <- unique(edge_)
 for (i in seq(1, nrow(edge_))) {
 n1 <- edge_[i, 1]
 n2 <- edge_[i, 2]
 neighbours[[n1]] <- c(n2, neighbours[[n1]])
 neighbours[[n2]] <- c(n1, neighbours[[n2]])
 }
 }
 structure(neighbours, class = "graph")
}
```

```{r}
profvis::profvis({
 n <- 20000
 nodes <- 1:n
 edges <- sample(nodes, 50000, replace = TRUE)
 weights <- rnorm(n)
 g <- graph(n, edges)
 smooth_weights(g, weights, 0.8)
})
```

```{r}
flow_weights_iteration <- function(graph, node_weights, alpha) {
 if (length(node_weights) != length(graph))
 stop("Incorrect number of nodes")
   no_nodes <- length(node_weights)
 new_weights <- vector("numeric", n)
 for (i in 1:no_nodes) {
 neighbour_weights <- 0
 n <- 0
 for (j in graph[[i]]) {
 if (i != j) {
 neighbour_weights <- neighbour_weights + node_weights[j]
 n <- n + 1
 }
 }
 if (n > 0) {
 new_weights[i] <- (alpha * node_weights[i] + (1 - alpha)
 * neighbour_weights / n)
 } else {
 new_weights[i] <- node_weights[i]
 }
 }
 new_weights
}
smooth_weights <- function(graph, node_weights, alpha, no_iterations) {
 new_weights <- node_weights
 replicate(no_iterations, {
 new_weights <- flow_weights_iteration(graph, new_weights, alpha)
 })
 new_weights
}
```

```{r message=FALSE}
profvis::profvis({
 n <- 20000
 nodes <- 1:n
 edges <- sample(nodes, 100000, replace = TRUE)
 weights <- rnorm(n)
 g <- graph(n, edges)
 smooth_weights(g, weights, 0.8, 10)
})
```

```{r}
smooth_weights <- function(graph, node_weights,
 alpha, no_iterations) {
 new_weights <- node_weights
 for (i in 1:no_iterations) {
 new_weights <-
 smooth_weights_iteration(graph, new_weights, alpha)
 }
 new_weights
}
```

```{r}
### Increasing Speed
library(microbenchmark)

mysum <- function(sequence) {
 s <- 0
 for (x in sequence) s <- s + x
 s
}
microbenchmark(
 sum(1:10),
 mysum(1:10)
)

microbenchmark(
 sum(1:10),
 mysum(1:10),
 Reduce(`+`, 1:10)
)
```

```{r}
x <- sample(LETTERS, 1000, replace = TRUE)
microbenchmark(
 factor(x, levels = LETTERS),
 factor(x)
)
```

```{r}
x <- rnorm(1000)
names(x) <- paste("n", 1:1000)
microbenchmark(
 unlist(Map(function(x) x**2, x), use.names = FALSE),
 unlist(Map(function(x) x**2, x))
)
```

```{r}
smooth_weights_iteration_map <- function(graph, node_weights, alpha) {
 if (length(node_weights) != length(graph))
 stop("Incorrect number of nodes")
handle_i <- function(i) {
 neighbour_weights <- 0
 n <- 0
 for (j in graph[[i]]) {
 if (i != j) {
 neighbour_weights <- neighbour_weights + node_weights[j]
 n <- n + 1
 }
 }
 if (n > 0) {
 alpha * node_weights[i] + (1 - alpha) * neighbour_weights / n
 } else {
 node_weights[i]
 }
 }
 unlist(Map(handle_i, 1:length(node_weights)))
}
```

library(cluster) test_rmse \<- function(data) { model \<-
data$training %>% lm(dist ~ speed, data = .)
 predictions <- data$test %\>% predict(model, data = .)
rmse(data$test$dist, predictions) } sample_rmse \<- function (n) {
random_cars \<- cars %\>% partition(n, c(training = 0.6, test = 0.4))
unlist(Map(test_rmse, random_cars)) } sample_rmse_parallel \<- function
(n) { random_cars \<- cars %\>% partition(n, c(training = 0.6, test =
0.4)) unlist(clustermq(cl, test_rmse, random_cars)) }

microbenchmark( sample_rmse(10), sample_rmse_parallel(10), times = 5 )

microbenchmark( sample_rmse(1000), sample_rmse_parallel(1000), times = 5
)

```{r}
library(Rcpp)
cppFunction("
NumericVector
smooth_weights_iteration_cpp(List g,
 NumericVector node_weights,
double alpha)
{
 NumericVector new_weights(g.length());
 for (int i = 0; i < g.length(); ++i) {
 // The body here is just the C++ code
 // shown above...
 }
 return new_weights;
}
")
```

```{r}
smooth_weights_cpp <- function(graph, node_weights,
 alpha, no_iterations) {
 new_weights <- node_weights
 for (i in 1:no_iterations) {
new_weights <-
 smooth_weights_iteration_cpp(graph, new_weights, alpha)
 }
 new_weights
}
```

library(microbenchmark) microbenchmark( smooth_weights(g, weights, 0.8,
10), smooth_weights_cpp(g, weights, 0.8, 10), times = 5 )

```{r}
library(rstudioapi)

viewer <- getOption("viewer")
viewer("R-with Data Science.Rmd", height = "maximize")
#rstudioapi::navigateToFile("")
```